### 2012-05-22 paper/section-background.tex

\section{Background and Related Work}
\label{sec:background}