### 2012-05-29 paper/section-background.tex
Add AUCTeX commands to deal with multifile document

\section{Background and Related Work}
\label{sec:background}

%%% Local Variables:
%%% TeX-master: "paper"
%%% End: