### 2012-06-07 paper/section-background.tex
Some notes on background work

\section{Background and Related Work}
\label{sec:background}

\begin{enumerate}
\item Reynolds
\item Kennedy
\item Gallier (for information on geometry)
\item Tse and Zdancewic, and Sabelfeld and Sands, for information flow
\item Chaudhuri, Gulwani and Lublinerman, for continuity analysis
\end{enumerate}

%%% Local Variables:
%%% TeX-master: "paper"
%%% End: