### 2012-06-08 paper/geometry-twocolumn.tex
Latest changes

\section{Motivating Examples}
\label{sec:motivating-examples}

We motivate this paper's goal of investigating algebraically-indexed
types and their relational interpretations by presenting several
examples. Our main set of examples comprise a novel type system for
writing programs that manipulate geometric data, while maintaining
geometric invariants through the types. \fixme{and the information
  flow examples}

\subsection{2-Dimensional Geometry and Origin Invariance}

When writing programs that manipulate geometric data, the basic data
structure is the $n$-tuple of numbers. In the 2-dimensional case,
tuples $\vec{v} = (x,y)$ are called upon to represent both
\emph{points}---offsets from some origin---and
\emph{vectors}---offsets in their own right. Points and vectors are
very different despite their common representation.

Distinguishing between points and vectors is the key feature of
\emph{affine geometry} (see, for example, Gallier's book
\cite{gallier11geometric}, Chapter 2). Traditionally in computational
geometry libraries, the distinction between points and vectors has
either been left to the programmer to maintain by themselves, or has
been enforced through the use of separate abstract types for points
and vectors (one might say that the standard mathematical formulation
of affine spaces uses the abstract types approach). Here, we
investigate a more sophisticated approach using types indexed by
change of origin transformations, and think of the difference between
points and vectors as a change of data representation.

Putting this in concrete terms, if we have two origins, $(0,0)$ and
$(10,20)$, then the tuple $(1,1)$ with respect to the first origin and
$(11,21)$ with respect to the second origin represent the \emph{same}
point. It is merely an artifact of our representation of points as
pairs of numbers that these appear to be different. When we write
programs that manipulate data that represents points, we ought to
ensure that our programs are invariant with respect to changes in the
choice of origin. On the other hand, vectors are not intended to be
regarded as invariant under change of origin. Vectors represent
offsets, or movements, and the vector $(0,0)$ always represents the
zero offset.

Invariance under change of representation immediately recalls
Reynolds' fable concerning two professors teaching the theory of
complex numbers \cite{reynolds83types}. One professor chooses
rectangular coordinates ($x + iy$), while the other chooses polar
coordinates ($\alpha\cos\theta + i\alpha\sin\theta$). Happily, after
learning the basic operations on complex numbers in the two
representations, the two classes can interact because the theory of
complex numbers is invariant with respect to the choice of
representation. Reynolds formalises the idea of invariance with
respect to changes of representation as preservation of relations. If
a binary relation $R$ relates two representations of a concept---for
example the rectangular and polar representations of the same complex
number are related---then a program that manipulates complex numbers
at a level of abstraction above their specific representation should
preserve the relation $R$. For example, let $f$ be a program taking
complex numbers to complex numbers that is intended to respect
invariance under change of representation. Then if $c$ is a complex
number in rectangular form, and $c'$ is a complex number in polar form
such that $R$ relates them, which we write as $(c,c') \in R$, then it
should be the case that $(f(c), f(c')) \in R$.

We now apply Reynolds' relational approach to the geometric
setting. We define a family of binary relations on $\mathbb{R}^2$
which is indexed by changes of origin. Changes of origin are
represented by vectors in $\mathbb{R}^2$, and form a group under
addition. To emphasise the group structure, and the intended meaning
as translations, we write the collection of all change of origin
offsets as $\Transl_2$. By quantifying over all changes of origin we
will be able to ensure that our programs are invariant with respect to
the particular choice of origin. The members of the
$\Transl_2$-indexed family of binary relations $\{ R_{\vec{t}}
\subseteq \mathbb{R}^2 \times \mathbb{R}^2 \}_{\vec{t} \in \Transl_2}$
are defined as follows:
\begin{displaymath}
  R_{\vec{t}} = \{ (\vec{v}, \vec{v'}) \sepbar \vec{v'} = \vec{v} + \vec{t} \}
\end{displaymath}

Now consider a function $f$ that, for example, takes two tuples in
$\mathbb{R}^2$ and returns a single tuple in $\mathbb{R}^2$. We intend
that all these tuples represent points with respect to the same
origin, but that the function itself should be invariant with respect
to the choice of origin. We use Reynolds' idea of preservation of
relations to formalise this: for any $\vec{t} \in \Transl_2$ (i.e.~any
change of origin), the function $f$ should satisfy the following
statement:
\begin{equation}\label{eq:f-preserve-rel-frame}
  \forall (\vec{v_1},\vec{v'_1}) \in R_{\vec{t}}, (\vec{v_2},\vec{v'_2}) \in R_{\vec{t}}. (f(\vec{v_1}, \vec{v_2}), f(\vec{v'_1}, \vec{v'_2})) \in R_{\vec{t}}
\end{equation}
Unfolding the definition of $R_{\vec{t}}$, this statement is
equivalent to the following statement expressing invariance of $f$
with respect to changes of origin: for all $\vec{t} \in \Transl_2$:
\begin{equation}\label{eq:f-invariant-frame}
  \forall \vec{v_1}, \vec{v_2}.\ f(\vec{v_1} + \vec{t},\vec{v_2} + \vec{t}) = f(\vec{v_1},\vec{v_2}) + \vec{t}.
\end{equation}
So Reynolds' preservation of relations, when instantiated with the
family of relations $\{R_{\vec{t}}\}$, yields exactly the geometric
property of invariance under change of origin.

% An example function $f$ that satisfies
% \statementref{eq:f-invariant-frame} is the following function that
% computes a particular affine combination of two points by working
% directly on their coordinate representation:
% \begin{displaymath}
%   f(\vec{v_1}, \vec{v_2}) = \frac{1}{2}\vec{v_1} + \frac{1}{2}\vec{v_2}.
% \end{displaymath}
% In general, affine combinations of points $\lambda_1\vec{v_1} +
% \lambda_2\vec{v_2}$, where $\lambda_1 + \lambda_2 = 1$, satisfy
% \statementref{eq:f-invariant-frame}. Affine combination is one of the
% fundamental building blocks of affine geometry -- the properties of
% points invariant under invertible affine maps
% (\cite{gallier11geometric}, Chapter 2). If we drop the condition that
% $\lambda_1 + \lambda_2 = 1$, then we are dealing with linear
% combinations of vectors, and we are no longer invariant with respect
% to changes of frame. However, linear combinations are invariant with
% respect to change of basis. We can represent changes of basis as
% linear invertible maps $B : \mathbb{R}^2 \to \mathbb{R}^2$. The set of
% all such maps forms the general linear group $\GL(2)$. We now define
% another family of relations $\{R_{\texttt{vec}}(B) \subseteq
% \mathbb{R}^2 \times \mathbb{R}^2 \}_{B \in \GL(2)}$ that relates two
% points up to change of basis:
% \begin{displaymath}
%   R_{\texttt{vec}}(B) = \{ (\vec{v_1},\vec{v_2}) \sepbar B\vec{v_2} = \vec{v_1} \}.
% \end{displaymath}
% Now the functions $f_{\lambda_1\lambda_2}(\vec{v_1},\vec{v_2}) =
% \lambda_1\vec{v_1} + \lambda_2\vec{v_2}$, for arbitrary $\lambda_1$
% and $\lambda_2$, do preserve the relations $R_{\texttt{vec}}(B)$, for
% all $B \in \GL(2)$. Unfolding the definition of $R_{\texttt{vec}}(B)$
% in the analogous statement to \statementref{eq:f-preserve-rel-frame}
% for $R_{\texttt{vec}}$ instead of $R_{\texttt{pt}}$, we can see that
% preservation of the relations $R_{\texttt{vec}}$ characterises the
% functions that are invariant under change of basis. For all $B \in \GL(2)$, we have
% \begin{displaymath}
%   \forall \vec{v_1}, \vec{v_2}.\ f(B\vec{v_1},B\vec{v_2}) = B(f(\vec{v_1},\vec{v_2})),
% \end{displaymath}
% and this is exactly the property of invariance under change of basis
% we required above of programs manipulating vectors.

% Note that the family of relations $R_{\texttt{vec}}$ is just
% $R_{\texttt{pt}}$ when restricted to elements of the group
% $\GL(2)$. In the type system we introduce in the next section, we
% combine points and vectors into the same data type. Whether it
% represents a point or a vector depends on the group of geometric
% transformations that we expect it to be invariant under.

% FIXME: vectors are not invariant under change of origin, so they are
% represented by the relation $R_0$, which is just equality. But they,
% and points are invariant under change of basis. However, as we shall
% see below, not all operations are invariant under all changes of
% basis. In particular the dot product is only invariant under
% orthogonal transformations.

\subsection{A Type System for Origin Invariance}
\label{sec:type-system-geom-intro}

Reynolds' further insight was to see how preservation of relations by
the denotational interpretations of programs can automatically be
established by following a typing discipline. In Reynolds' case, the
typing discipline was the polymorphic $\lambda$-calculus where new
types can be constructed by universal quantification over all
types. Universal type quantification is commonly written $\forall
\alpha. A$, which denotes a type that may be instantiated with any
other type $B$ by replacing the type variable $\alpha$ with $B$ in
$A$.

In terms of relations, Reynolds interprets universal quantification
over types as quantification over binary relations between denotations
of types. In the statements of geometric invariance that we stated in
the previous section (e.g., \statementref{eq:f-invariant-frame}), we
did not quantify over all relations. Instead, we quantified over all
changes of origin and used the choice to select a relation from the
family $\{R_{\vec{t}}\}$. Driven by this observation, we introduce
quantification over changes of origin into the language of types.  We
use the notation $\forall t \mathord: \SynTransl_2. A$ for
quantification over all 2-dimensional translations (i.e.~choices of
origin) $t$. We refer to $\SynTransl_2$ as the \emph{sort} of
$t$. Note that we have used a different font to distinguish the
semantic group $\Transl_2$ from the syntactic sort $\SynTransl_2$.

The sort $\SynTransl_2$ represents an abelian group, so we can combine
its elements using the usual group operations. We write operations
additively, using $e_1 + e_2$ for the group operation, $-e$ for
inverse and $0$ for the unit.  We also regard expressions built from
variables and the group operations up to the abelian group axioms. For
example, $e_1 + (e_2 + e_3)$ and $(e_1 + e_2) + e_3$ are to be
regarded as equivalent.

Our language of types also includes the common type constructors for
function types, $A \to B$, sum types $A + B$, unit type
$\tyPrimNm{unit}$ and tuple types $A \times B$. We also assume a
primitive type $\tyPrimNm{real}$, used to represent scalars. As we
noted at the start of the previous section, the central data structure
in geometric applications is the tuple of real numbers for
representing points and vectors. However, we cannot simply express
this as the type $\tyPrimNm{real} \times \tyPrimNm{real}$ because it
will not have the correct relational interpretation. (Two elements of
type $\tyPrimNm{real}$ will be related if and only if they are equal,
and hence two elements of $\tyPrimNm{real} \times \tyPrimNm{real}$
will be related if and only if they are equal, by Reynolds' definition
of the relational interpretation of tuple types.) Therefore, we
introduce a new type parameterised by expressions $e$ of sort
$\SynTransl_2$ to represent points \emph{and} vectors:
$\tyPrim{vec}{e}$. We have used the name $\tyPrimNm{vec}$, even though
we have taken pains to separate geometric points and vectors, to
recall the computer science notion of vector as a sequence of values
of homogeneous type which has a known length (in this case two).

We now sketch the interpretation of the types $\tyPrim{vec}{e}$ and
$\forall t\mathord:\SynTransl_2.A$. In our calculus, each type has two
interpretations: an index-erasure interpretation that ignores the
indexing expression $e$, and a relational interpretation as a binary
relation on the index-erasure interpretation. For now, we assume we
are given some way of turning an expression $e$ of sort $\SynTransl_2$
into an actual element $\sem{e}\rho$ of the group $\Transl_2$, under
some environment $\rho$ which tells how to interpret the free
variables. The (simplified) index-erasure and relation interpretations
of the types $\tyPrim{vec}{e}$ are:
\begin{displaymath}
  \begin{array}{l@{\hspace{0.5em}=\hspace{0.5em}}l}
    \tySem{\tyPrim{vec}{e}} & \mathbb{R}^2 \\
    \rsem{\tyPrim{vec}{e}}{}\rho & R_{\sem{t}\rho} = \{ (\vec{v},\vec{v'}) \sepbar \vec{v'} = \vec{v} + \sem{e}\rho \}
  \end{array}
\end{displaymath}
The index-erasure interpetation of universal quantification $\forall
t\mathord:\SynTransl_2.A$ is simply the interpretation of the
underlying type $A$. The relational interpretation intersects the
relational interpretations of the type $A$ under all extensions of the
relational environment $\rho$:
\begin{displaymath}
  \begin{array}{l@{\hspace{0.5em}=\hspace{0.5em}}l}
    \tySem{\forall t\mathord:\SynTransl_2.A} & \tySem{A} \\
    \rsem{\forall t\mathord:\SynTransl_2.A}{}\rho & \bigcap\{ \rsem{A}{}{\rho'} \sepbar \rho' \in \extends{\rho}{i\mathord:\SynTransl_2} \}
  \end{array}
\end{displaymath}
The exact definition of $\extends{\rho}{i\mathord:\SynTransl_2}$
depends on which (possibly non-compositional) interpretation of
relational environments that we take. For now, it will suffice to
think of relational environments as simply mappings from free
variables to the group $\Transl_2$, and extension as just adding an
extra component to the tuple. The general index-erasure semantics will
be given in \autoref{sec:index-erasure-semantics}. The full relational
interpretation is in general more complex, due to the possibility of
non-compositional interpretations of free index variables. The
relational interpretation will be presented in
\autoref{sec:relational-semantics}.

Equipped with quantification over the group $\SynTransl_2$ of changes
of origin, and the indexed types $\tyPrim{vec}{\cdot}$, we can now
express origin invariance properties as types. In the previous
section, we discussed functions $f : \mathbb{R}^2 \times \mathbb{R}^2
\to \mathbb{R}^2$ that have the property that they preserve all
changes of origin. We can now express this property as a type:
\begin{displaymath}
  \mathrm{f} : \forall t \mathord: \SynTransl_2.\ \tyPrim{vec}{t} \times \tyPrim{vec}{t} \to \tyPrim{vec}{t}
\end{displaymath}
Spelling out the relational interpretation of this type using the
definitions above (and the standard relational interpretations for
tuple and function types), we obtain
\statementref{eq:f-invariant-frame}.

\subsection{Affine and Vector Operations}
\label{sec:affine-vector-ops}

Invariance under change of origin is the key feature of affine
geometry, and the central operation of affine geometry is the affine
combination of points: $\lambda_1\vec{v_1} + \lambda_2\vec{v_2}$,
where $\lambda_1 + \lambda_2 = 1$. Geometrically, affine combination
can be interpreted as interpolation between the points represented by
$\vec{v_1}$ and $\vec{v_2}$. We add affine combination of points to
our calculus, with the following typing and intended denotation:
\begin{displaymath}
  \begin{array}{l}
    \mathrm{affComb} :\forall t \mathord: \SynTransl_2.\ \tyPrim{vec}{t} \to \tyPrimNm{real} \to \tyPrim{vec}{t} \to \tyPrim{vec}{t} \\
    \sem{\mathrm{affComb}}\ \vec{v_1}\ r\ \vec{v_2} = (1-r)\vec{v_1} + r\vec{v_2}
  \end{array}
\end{displaymath}
It can be verified by hand that the intended denotation
$\sem{\mathrm{affComb}}$ has the invariance property specified by its
type.
% By defining the primitive function $\mathrm{affComb}$ to take a
% single $\tyPrimNm{real}$ parameter $t$, we can easily ensure that we
% are taking the affine combination of two representatives of points,
% and not just an arbitrary linear combination.

\begin{example}
  The evaluation of quadratic B\'{e}zier curves (B\'{e}zier curves
  with two endpoints and a single control point) can be expressed
  using the affine combination primitive and three steps of De
  Casteljau's algorithm:
  \begin{displaymath}
    \begin{array}{@{}l}
      \mathrm{quadB\acute{e}zier} : \forall t \mathord:\SynTransl_2.\ \tyPrim{vec}{t} \to \tyPrim{vec}{t} \to \tyPrim{vec}{t} \to \tyPrimNm{real} \to \tyPrim{vec}{t} \\
      \mathrm{quadB\acute{e}zier}\ [t]\ p_0\ p_1\ p_2\ s = \\
      \quad \mathrm{affComb}\ [t]\ (\mathrm{affComb}\ [t]\ p_0\ s\ p_1)\ s\ (\mathrm{affComb}\ [t]\ p_1\ s\ p_2) \\
    \end{array}
  \end{displaymath}
  For two endpoints $p_0$ and $p_2$ and a control point $p_1$, an
  application $\mathrm{quadB\acute{e}zier}\ p_0\ p_1\ p_2\ s$ for $s
  \in [0,1]$ gives us the point on the curve at ``time'' $s$. Note
  that the type of $\mathrm{quadB\acute{e}zier}$ tells us immediately
  that it preserves all changes of origin.
\end{example}

% Affine combination only provides us with an operation on points. We
% are also interested in operations on the vectors representing offsets
% between points. We now examine the correct types to assign to the
% vector space operations of addition of vectors, negation of vectors,
% multiplication by a scalar and the zero vector. The typings of these
% operations will make use of the abelian group structure of change of
% origin translations.

\fixme{Need to make clear that the point is to make use of the
  algebraic indexing structure}

Vector operations are not invariant under change of origin. The
obvious type for vector addition is therefore:
\begin{displaymath}
  (+) : \tyPrim{vec}{0} \to \tyPrim{vec}{0} \to \tyPrim{vec}{0}
\end{displaymath}
(we write binary operators intended to be used infix in parentheses
when not appearing in infix position, following the Haskell
syntax.)

However, it is possible to give vector addition a more precise type
that describes its effect on the chosen origin. 
\begin{displaymath}
  (+) : \forall t_1, t_2 \mathord: \SynTransl_2.\ \tyPrim{vec}{t_1} \to \tyPrim{vec}{t_2} \to \tyPrim{vec}{t_1 + t_2}
\end{displaymath}
We can also negate vectors, yielding a vector which points in the
opposite direction. Negation negates translation arguments:
\begin{displaymath}
  \mathrm{negate} : \forall t \mathord: \SynTransl_2.\ \tyPrim{vec}{t} \to \tyPrim{vec}{-t}
\end{displaymath}
With the primitive operations of addition and negation of vectors, we
can define the derived operation of subtraction:
\begin{displaymath}
  \begin{array}{l}
    (-) : \forall t_1,t_2 \mathord:\SynTransl_2.\ \tyPrim{vec}{t_1} \to \tyPrim{vec}{t_2} \to \tyPrim{vec}{t_1 - t_2} \\
    (-)\ [t_1]\ [t_2]\ p_1\ p_2 = p_1 + \mathrm{negate}\ p_2
  \end{array}
\end{displaymath}

Given two points that have been constrained to be invariant with
respect to the same change of origin---i.e.~two values of type
$\tyPrim{vec}{t}$---we can compute their offset, which is a vector
expressed with respect to the null change of origin. The offset
operation is a special case of vector subtraction:
\begin{displaymath}
  \begin{array}{l}
    \mathrm{offset} : \forall t \mathord:\SynTransl_2.\ \tyPrim{vec}{t} \to \tyPrim{vec}{t} \to \tyPrim{vec}{0} \\
    \mathrm{offset}\ [t]\ p_1\ p_2 = p_1 - p_2
  \end{array}
\end{displaymath}
The type of the vector addition operation can be specialised to the
case of moving a point by a vector:
\begin{displaymath}
  \begin{array}{l}
    \mathrm{moveBy} : \forall t \mathord:\SynTransl_2.\ \tyPrim{vec}{t} \to \tyPrim{vec}{0} \to \tyPrim{vec}{t} \\
    \mathrm{moveBy}\ [t]\ p\ v = p + v
  \end{array}
\end{displaymath}
The types we assign to the remaining vector space primitives, the zero
vector and multiplication by a scalar, do not describe any interesting
effect on translations:
\begin{displaymath}
  \begin{array}{l@{\hspace{0.5em}:\hspace{0.5em}}l}
    \mathrm{0} & \tyPrim{vec}{0}\\
    (*) & \tyPrimNm{real} \to \tyPrim{vec}{0} \to \tyPrim{vec}{0}
  \end{array}
\end{displaymath}

% \fixme{If we had dependent types, could the affComb operation be
%   derived from an appropriately dependently typed scaling operation?}

\begin{example}
  The vector space operators and the properties that follow from their
  types allow us to establish a useful type isomorphism. Consider
  functions with types following the schema:
  \begin{displaymath}
    \tau_{n} \isDefinedAs \forall t\mathord:\SynTransl_2.\ \underbrace{\tyPrim{vec}{t} \to ... \tyPrim{vec}{t}}_{n+1\textrm{ times}} \to \tyPrimNm{real}
  \end{displaymath}
  Just by looking at the types $\tau_{n}$, we know that their
  inhabitants will be invariant under change of origin, due to the
  quantification over all $t$ in $\SynTransl_2$. Therefore, we may as well pick
  one of the input points as the origin and assume that all the other
  points are defined with respect to this chosen origin. This is
  similar to the common mathematical step of stating that, ``without
  loss of generality'', we may pick some point in a description of a
  problem to be the origin, as long as the problem statement is
  invariant under translation.

  The types $\tau_{n}$ are isomorphic to the corresponding types
  $\sigma_{n}$:
  \begin{displaymath}
    \sigma_{n} \isDefinedAs \underbrace{\tyPrim{vec}{0} \to ... \tyPrim{vec}{0}}_{n\textrm{ times}} \to \tyPrimNm{real}
  \end{displaymath}
  We demonstrate this isomorphism formally in
  \autoref{sec:instantiations}, in a more general setting of types
  indexed by abelian groups.

  % The isomorphism between these two families of types can be witnessed
  % by the following two functions.
  % \begin{displaymath}
  %   \begin{array}{l}
  %     i_{n,A} : \tau_{n,A} \to \sigma_{n,A} \\
  %     i_{n,A}\ f\ [B]\ (p_1, p_2, ..., p_n) = f\ [B \ltimes 0]\ (0, p_1, ..., p_n) \\
  %     \\
  %     i^{-1}_{n,A} : \sigma_{n,A} \to \tau_{n,A} \\
  %     i^{-1}_{n,A}\ g\ [F]\ (p_0, ..., p_n) = g\ [\pi_L(F)]\ (p_1-p_0, ..., p_n-p_0)
  %   \end{array}
  % \end{displaymath}
  % The direction defined by the function $i_{n,A}$ treats the inputs
  % $p_1,...,p_n$ as vector offsets from the origin $0$. The direction
  % witnessed by $i^{-1}_{n,A}$ picks the first point to act as the
  % origin, and uses the operation of vector subtraction we defined
  % above to turn each of the other points into offsets from this
  % point. Note that this isomorphism is not unique: for each $n$ we can
  % pick any of the $n+1$ inputs to act as the distinguished
  % origin. Thus the families of types $\tau_{n,A}$ and $\sigma_{n,A}$
  % are isomorphic, but not uniquely isomorphic.
\end{example}

\begin{example}
  An inexpressible type
  \begin{displaymath}
    \forall t \mathord: \SynTransl_2.\ \tyPrim{vec}{t + t} \to \tyPrim{vec}{t}
  \end{displaymath}
  FIXME: explain that it is possible to show that this type is
  uninhabited with the given operations.
\end{example}

\subsection{Change of Basis Invariance}
\label{sec:motivation-generalising}

From basic linear algebra, we know that the vector operations of
addition, negation and scaling, while not invariant under change of
origin, are invariant under change of basis. Following the same
methodology as we did for origin invariance, we can express basis
invariance by means of preservation of relations indexed by changes of
basis. Change of basis is accomplished by applying an invertible
linear map, and the collection of all invertible linear maps on
$\mathbb{R}^2$ forms the group $\GL_2$.

\fixme{new relational interpretation of vectors}

\begin{displaymath}
  \begin{array}{l@{\hspace{0.5em}=\hspace{0.5em}}l}
    \tySem{\tyPrim{vec}{e_B,e_t}} & \mathbb{R}^2 \\
    \rsem{\tyPrim{vec}{e_B,e_t}}{}\rho & \{ (\vec{v},\vec{v'}) \sepbar \vec{v'} = \sem{e_B}\rho\vec{v} + \sem{e_t}\rho \}
  \end{array}
\end{displaymath}

\paragraph{Affine Geometry} The combination of a change of basis
transformation and a change of origin transformation is known as an
\emph{affine transformation}. A way of describing affine geometry is
as the collection of properties that are invariant under invertible
affine transformations.

\begin{eqnarray*}
  \mathrm{affComb} & : &
  \begin{array}[t]{@{}l}
    \forall B \mathord: \SynGL_2, t \mathord: \SynTransl_2.\\
    \hspace{0.2cm} \tyPrim{vec}{B,t} \to \tyPrimNm{real} \to \tyPrim{vec}{B,t} \to \tyPrim{vec}{B,t}
  \end{array}
  \\
  (+) & : &
  \begin{array}[t]{@{}l}
    \forall B \mathord: \SynGL_2, t_1,t_2 \mathord: \SynTransl_2. \\
    \hspace{0.2cm}\tyPrim{vec}{B,t_1} \to \tyPrim{vec}{B,t_2} \to \tyPrim{vec}{B,t_1 + t_2}
  \end{array}
  \\
  \mathrm{negate} & : & \forall B \mathord: \SynGL_2, t \mathord: \SynTransl_2.\ \tyPrim{vec}{B,t} \to \tyPrim{vec}{B,-t} \\
  0 & : & \forall B \mathord: \SynGL_2.\ \tyPrim{vec}{B,0} \\
  (*) & : & \forall B \mathord: \SynGL_2.\ \tyPrimNm{real} \to \tyPrim{vec}{B,0} \to \tyPrim{vec}{B,0}
\end{eqnarray*}

\fixme{we now have all the operations of affine geometry, along with
  the property that they are invariant under affine transformations.}

\paragraph{Euclidean Geometry} Euclidean geometry extends affine
geometry with the operation of \emph{dot product} (also known as the
\emph{inner product}) of two vectors, which produces a real
number. The standard Euclidean dot product is defined on the
coordinate representation in the following way:
\begin{displaymath}
  (x_1,y_1) \cdot (x_2,y_2) = x_1x_2 + y_1y_2
\end{displaymath}
% The dot product is used to define properties of vectors such as their
% \emph{norm} $||\vec{v}|| = \vec{v}\cdot\vec{v}$, which is the square
% of the length of the vector $\vec{v}$, and the notion of
% orthogonality: two vectors $\vec{v_1}$ and $\vec{v_2}$ are orthogonal
% if $\vec{v_1}\cdot\vec{v_2} = 0$.

To incorporate the operation of dot product into our calculus, we must
give it a type. It is not the case that the dot product is invariant
under the action of either of the groups $\GL_2$ or $\Transl_2$ that
we have considered so far. In fact, the dot product is invariant under
the subgroup $\Orth_2$ of $\GL_2$ of \emph{orthogonal} linear
transformations. The group $\Orth_2$ consists of invertible linear
maps whose determinant is either $1$ or $-1$. We introduce orthogonal
transformations as a new sort $\SynOrth_2$, and overload the
multiplicative group operations syntax for inhabitants of
$\SynOrth_2$. We also assume a function $\iota_O$ that takes $e :
\SynOrth_2$ to $\iota_O(e) : \SynGL_2$. We may now assign the following type
to the dot product:
\begin{displaymath}
  (\cdot) : \forall O \mathord: \SynOrth(2).\ \tyPrim{vec}{\iota_OO, 0} \to \tyPrim{vec}{\iota_OO, 0} \to \tyPrimNm{real}
\end{displaymath}

\begin{figure}[t]
  \centering
  \begin{eqnarray*}
    0   &:& \forall s \mathord:\SynGL_1.\ \tyPrim{real}{s} \\
    1   &:& \tyPrim{real}{1} \\
    (+) &:& \forall s \mathord:\SynGL_1.\ \tyPrim{real}{s} \to \tyPrim{real}{s} \to \tyPrim{real}{s} \\
    (-) &:& \forall s \mathord:\SynGL_1.\ \tyPrim{real}{s} \to \tyPrim{real}{s} \to \tyPrim{real}{s} \\
    (*) &:& \forall s_1,s_2 \mathord:\SynGL_1.\ \tyPrim{real}{s_1} \to \tyPrim{real}{s_2} \to \tyPrim{real}{s_1s_2} \\
    (/) &:&
    \begin{array}[t]{@{}l@{}l}
      \forall s_1,s_2 \mathord:\SynGL_1.\ \tyPrim{real}{s_1}\ & \to \tyPrim{real}{s_2} \\
      &\to \tyPrim{real}{s_1s_2^{-1}} + \tyPrimNm{unit} \\
    \end{array}\\
    \mathrm{abs} &:& \forall s \mathord:\SynGL_1.\ \tyPrim{real}{s} \to \tyPrim{real}{|s|} %\\
%    \mathrm{sqrt} &:& \tyPrim{real}{1} \to \tyPrim{real}{1}
  \end{eqnarray*}
  \caption{Operations on scaled real numbers}
  \label{fig:real-ops}
\end{figure}

The cross product of two vectors is defined on coordinate
representations as follows:
\begin{displaymath}
  (x_1,y_1) \times (x_2,y_2) = x_1y_2 - x_2y_1
\end{displaymath}
Geometrically, the cross product is the signed area of the
parallelogram described by the pair of input vectors. It turns out
that the value of the cross product of two vectors varies with the
determinant of the matrix describing the basis of the vector space. We
therefore augment our calculus with a new sort $\SynGL_1$ of scale
factors. Semantically, $\SynGL_1$ ranges over the non-zero real
numbers and forms an abelian group, which we write
multiplicatively. We also add two new operations: determinant, $\det
B$, which takes an inhabitant of $\SynGL_2$ to its determinant in
$\SynGL_1$, and the operation of absolute value that takes scaling
factors to scaling factors which we write as $|e|$. We also refine the
type $\tyPrimNm{real}$ of real numbers to take a parameter of sort
$\SynGL_1$: $\tyPrim{real}{e}$, where the old type $\tyPrimNm{real}$
is just $\tyPrim{real}{1}$. The collection of operations on real
numbers with scaling factors is shown in \autoref{fig:real-ops}.

With the additional sort $\SynGL_1$, and the refined type of real
numbers, we can state the type of the cross product.
\begin{displaymath}
  (\times) : \forall B \mathord: \SynGL_2.\ \tyPrim{vec}{B,0} \to \tyPrim{vec}{B,0} \to \tyPrim{real}{\det B}
\end{displaymath}
For orthogonal transformations $O : \Orth_2$, the absolute value of
the determinant is always $1$, so we assume the equation $|\det
(\iota_O O)| = 1$ to hold.

\begin{example}
  Using the structure we have defined in this subsection, we can
  define the following function that computes the area of a
  triangle. By the type of this function, we can see that the area of
  a triangle is invariant under orthogonal transformations and
  arbitrary translations \fixme{i.e. isometries}.
  \begin{displaymath}
    \begin{array}{@{}l}
      \mathrm{area} : \forall O\mathord:\SynOrth_2, t\mathord:\SynTransl_2.\ \\
      \hspace{0.8cm}\tyPrim{vec}{\iota_OO, t} \to \tyPrim{vec}{\iota_OO, t} \to \tyPrim{vec}{\iota_OO, t} \to \tyPrim{real}{1} \\
      \mathrm{area}\ [O]\ [t]\ p_1\ p_2\ p_3 = \frac{1}{2} * \mathrm{abs}\ ((p_2 - p_1) \times (p_3 - p_1))
    \end{array}
  \end{displaymath}
  In the body of this function, we have performed the calculation in
  several steps, each of which removes some of the symmetry described
  by the type of $\mathrm{area}$. First we have computed the two
  offset vectors $p_2 - p_1$ and $p_3 - p_1$. These operations remove
  the effect of translations on the result, in exactly the same way as
  the type isomorphism we demonstrated above \fixme{ref}. We then
  compute the cross product of the two vectors to determine the area
  of the parallelogram described by the sides of the triangle. This
  value has type $\tyPrim{real}{\det (\iota_O O)}$. We know that the
  determinant of an orthogonal transformation is always either $-1$ or
  $1$, so we have a result which varies according to possible
  reflections of the space. To remove this variation, we apply the
  $\mathrm{abs}$ function to leave us with a value of type
  $\tyPrim{real}{|\det (\iota_O O)|}$. This type is equal to the type
  $\tyPrim{real}{1}$. Finally, we multiply by $\frac{1}{2}$ to recover
  the area of the triangle, rather than the whole parallelogram.
\end{example}

% \begin{enumerate}
% \item Basis invariance
% \item Orthogonal transform invariance for dot product
% \item Rotational invariance (the type of the cross-product)
% \item Scale invariance (generalisation of units of measure)
% \end{enumerate}

% In some cases, we can use the invariance as a programming tool
% (e.g. type isomorphisms), but at others we can only use it from the
% outside.

% \begin{enumerate}
% \item Possibly do a $n$-body gravity simulator as an example?
% \item Have a special sort and type for rotations. An operation for
%   computing the rotation between two points, and for applying a
%   rotation to a vector. This will allow for more free theorems...
% \end{enumerate}

\subsection{Scale Invariance and Dimension Analysis}
\label{sec:scale-invariance}

\begin{enumerate}
\item Explicitly link scale invariance to units of measure system
\end{enumerate}

Need a function $\iota_1 : \SynGL_1 \to \SynGL_2$. Such that $\det (\iota_1 s) = s^2$

\subsection{Information Flow}
\label{sec:information-flow}

Blah blah blah. Need to work this out in terms of representation indepdence.

\begin{enumerate}
\item Think of information flow as passing around unforgable tokens
  \begin{enumerate}
  \item Give an example (given a piece of data protected by a token
    and a token, can we extract the data?)
  \item Need to make sure that tokens are unforgable
  \end{enumerate}
\item Could either enforce unforgability of tokens by some dynamic
  means, or by static type-based means
  \begin{enumerate}
  \item Leave the dynamic method as vague
  \item Ingredients of the static method: run-time tokens are just
    dummy values ($*$), the token type is parameterised by logical
    formulae, and the relational meaning is derived from an assignment
    of boolean values to atoms.
  \item Explain this as invariance under changes of satisfaction of
    the participants. Relate to descriptions of logics as rules that
    don't care about truth, only its preservation.
  \end{enumerate}
\item Cite Sabelfeld and Sands: ``A PER Model of Secure Information
  Flow in Sequential Programs''. And also Tse and Zdancewic.
\item Go a step further by considering equivalence relations on
  word32, and their lattice structure. Link this to quantative
  information flow. Will also need to include singleton types.
\end{enumerate}

\subsection{Basic Continuity Analysis}
\label{sec:continuity-analysis}

Doing epsilon-delta proofs!

Have a new sort $\mathsf{R}^{>0}$, representing positive distances
between real numbers.

\begin{displaymath}
  \begin{array}{l@{\hspace{0.5em}=\hspace{0.5em}}l}
    \tySem{\tyPrim{real}{e}} & \mathbb{R} \\
    \rsem{\tyPrim{real}{e}}{}\rho & \{ (x,x') \sepbar |x-x'| < \sem{e}\rho \}
  \end{array}
\end{displaymath}

A continuous function has type:
\begin{displaymath}
  \forall \epsilon \mathord: \mathsf{R}^{>0}.\ \exists \delta\mathord: \mathsf{R}^{>0}.\ \tyPrim{real}{\delta} \to \tyPrim{real}{\epsilon}
\end{displaymath}

\begin{enumerate}
\item Need to give types to basic operations on real numbers
\item Need to add existential types to the calculus
\end{enumerate}

%%% Local Variables:
%%% TeX-master: "paper"
%%% End: