### 2012-06-12 paper/section-background.tex

\section{Background and Related Work}
\label{sec:background}

\begin{enumerate}
\item Reynolds \cite{reynolds83types}
\item Kennedy \cite{kennedy97relational}
\item Gallier (for information on geometry) \cite{gallier11geometric}
\item Tse and Zdancewic, and Sabelfeld and Sands, for information flow \cite{sabelfeld01per}
\item Chaudhuri, Gulwani and Lublinerman, for continuity analysis \cite{chaudhuri10continuity}
\end{enumerate}

We discuss further related work in \autoref{sec:discussion}.

%%% Local Variables:
%%% TeX-master: "paper"
%%% End: