### 2012-06-12 paper/section-discussion.tex

\section{Discussion}
\label{sec:discussion}

\begin{enumerate}
\item Dependent Types (allow for more precision, in some cases would
  subsume our general framework, reference Bernardy's work)
\item Type-indexed types (e.g.~type and effect systems)
\item Making use of the free theorems, either by internalisation or by
  deep embedding of our general calculus in a theorem prover.
\item Extension of the definability work. Is it possible to extend
  standard definability results for the $\lambda$-calculus to cover
  algebraically indexed types?
\item Discussion of the relationship between relational parametricity
  and the abstraction theorem we have demonstrated. Discussion of the
  identity extension property in this setting.
\item Extending the geometric example with linear transformations, or
  even higher-degree tensors.
\end{enumerate}

%%% Local Variables:
%%% TeX-master: "paper"
%%% End: