### 2012-06-14 paper/paper.tex

\documentclass[natbib,preprint]{sigplanconf}

\usepackage[usenames]{color}
\definecolor{citationcolour}{rgb}{0,0.4,0.2}
\definecolor{linkcolour}{rgb}{0,0,0.8}
\usepackage{hyperref}
\hypersetup{colorlinks=true,
            urlcolor=linkcolour,
            linkcolor=linkcolour,
            citecolor=citationcolour,
            pdftitle=Relational Parametricity for Algebraically-Indexed Types,
            pdfauthor={Robert Atkey, Neil Ghani, Patricia Johann, Andrew Kennedy},
            pdfkeywords={}}  
\def\sectionautorefname{Section}
\def\subsectionautorefname{Section}
\def\subsubsectionautorefname{Section}

\title{Relational Parametricity for Algebraically Indexed Types}

\authorinfo{Robert Atkey\and Neil Ghani\\\and Patricia Johann}
           {University of Strathclyde}
           {\{Robert.Atkey,Neil.Ghani,Patricia.Johann\}@cis.strath.ac.uk}

\authorinfo{Andrew Kennedy}
           {Microsoft Research}
           {}

\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage[all]{xy}

\newcommand{\GL}[1]{\mathrm{GL}_#1}
\newcommand{\SynGL}[1]{\mathsf{GL}_#1}
\newcommand{\Orth}[1]{\mathrm{O}_#1}
\newcommand{\SynOrth}[1]{\mathsf{O}_#1}
\newcommand{\Transl}[1]{\mathrm{T}_#1}
\newcommand{\SynTransl}[1]{\mathsf{T}_#1}

% \newcommand{\GLtwo}{\mathrm{GL}_2}
% \newcommand{\SynGLtwo}{\mathsf{GL}_2}
% \newcommand{\GLone}{\mathrm{GL}_1}
% \newcommand{\SynGLone}{\mathsf{GL}_1}

\newcommand{\sepbar}{\mathrel|}

\newcommand{\Rel}{\mathrm{Rel}}
\newcommand{\relArrow}{\mathrel{\widehat\to}}
\newcommand{\relTimes}{\mathrel{\widehat\times}}
\newcommand{\relSum}{\mathrel{\widehat+}}

\newcommand{\setOfIntegers}{\mathbb{Z}}
\newcommand{\setOfBooleans}{\{\tmTT, \tmFF\}}

\newcommand{\id}{\mathrm{id}}

\newcommand{\SortSet}{\mathit{Sort}}
\newcommand{\IndexOpSet}{\mathit{IndexOp}}
\newcommand{\PrimTypeSet}{\mathit{PrimType}}
\newcommand{\IndexAxiomSet}{\mathit{IndexAx}}

\newcommand{\indexOp}[1]{\texttt{#1}}
\newcommand{\idxTms}[2]{\mathrm{IdxTm}(#1 \vdash #2)}
\newcommand{\idxTmAS}[1]{\mathrm{IdxTm}(#1)}

\newcommand{\tyInt}{\texttt{int}}
\newcommand{\tyBool}{\texttt{bool}}
\newcommand{\tyUnit}{\texttt{unit}}
\newcommand{\tyPrim}[2]{\texttt{#1}\langle #2 \rangle}
\newcommand{\tyPrimNm}[1]{\texttt{#1}}
\newcommand{\primTyArity}{\mathrm{tyArity}}
\newcommand{\indexOpArity}{\mathrm{opArity}}

\newcommand{\tyArr}{\to}
\newcommand{\tyProduct}{\times}
\newcommand{\tyX}[1]{\texttt{X}\langle #1 \rangle}
\newcommand{\isType}{\textsf{ type}}
\newcommand{\isCtxt}{\textsf{ ctxt}}
\newcommand{\Ty}{\textsf{Ty}}
\newcommand{\ty}{\textsf{ty}}

\newcommand{\tmTT}{\texttt{tt}}
\newcommand{\tmFF}{\texttt{ff}}

\newcommand{\relEnv}[1]{\mathcal{#1}}
\newcommand{\tySem}[1]{\llbracket #1 \rrbracket^{\mathcal{T}}}
\newcommand{\ctxtSem}[1]{\llbracket #1 \rrbracket^{\mathcal{C}}}
\newcommand{\tmSem}[1]{\llbracket #1 \rrbracket^{\mathit{tm}}}
\newcommand{\tyPrimSem}[1]{\llbracket #1 \rrbracket^{\mathcal{T}_0}}
\newcommand{\rsem}[3]{\llbracket #1 \rrbracket^{\mathcal{R}}_{#2}{#3}}
\newcommand{\extends}[2]{\mathsf{ext}(#1,#2)}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newcommand{\lemref}[1]{\hyperref[#1]{Lemma \ref*{#1}}}
\newcommand{\thmref}[1]{\hyperref[#1]{Theorem \ref*{#1}}}
\newcommand{\propref}[1]{\hyperref[#1]{Proposition \ref*{#1}}}
\newcommand{\corref}[1]{\hyperref[#1]{Corollary \ref*{#1}}}
\newcommand{\conref}[1]{\hyperref[#1]{Conjecture \ref*{#1}}}
\newcommand{\exref}[1]{\hyperref[#1]{Example \ref*{#1}}}
\newcommand{\statementref}[1]{\hyperref[#1]{Statement \ref*{#1}}}

\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\isDefinedAs}{\stackrel{\mathit{def}}=}

\newtheoremstyle{examplestyle}
  {\topsep}  % space above
  {\topsep}  % space below
  {\normalfont}% name of font to use in the body of the theorem
  {0em}% measure of space to indent
  {\bfseries}% name of head font
  {.}% punctuation between head and body
  {5pt plus 1pt minus 1pt}% space after theorem head; " " = normal interword space
  {}% Manually specify head
\theoremstyle{examplestyle}
\newtheorem{example}{Example}

\newtheoremstyle{restatementstyle}
  {\topsep}  % space above
  {\topsep}  % space below
  {\itshape}% name of font to use in the body of the theorem
  {0em}% measure of space to indent
  {\bfseries}% name of head font
  {.}% punctuation between head and body
  {5pt plus 1pt minus 1pt}% space after theorem head; " " = normal interword space
  {\thmname{#1} \thmnote{(Restatement of #3)}}% Manually specify head
\theoremstyle{restatementstyle}
\newtheorem{restateLemma}{Lemma}
\newtheorem{restateTheorem}{Theorem}

\newcommand{\fixme}[1]{\textbf{FIXME: #1}}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\category{D.1.1}{Programming techniques}{Applicative (functional)
  programming} \category{D.2.4}{Software Engineering}{Software/Program
  Verification} \category{D.3.3}{Programming Languages}{Language
  Constructs and Features---Data types and structures}

\terms
  Languages, Theory, Types

\keywords
  parametricity, dimension types, computational geometry

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{section-introduction.tex}
%\input{section-motivating-examples.tex}
\input{geometry-twocolumn.tex}
%\input{section-background.tex}
\input{section-general-framework.tex}
\input{section-instantiations.tex}
\input{section-discussion.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plainnat}
\bibliography{paper}

\newpage
\appendix

\input{appendix-proofs.tex}

\end{document}