### 2012-07-06 paper/section-monoid.tex
Metric space example, and splitting of sections into files
- fill out the metric space example, and correct 'continuity' to be 'uniform continuity'
  - still need to flesh out the connection with differential privacy
- split individual examples into separate files
- move the abelian group proofs into the appendix, pending fixing up
- tiny changes to the singleton types and logical types sections

\section{Monoid Indexed Types}
\label{sec:monoid-indexed-types}

\newcommand{\Mon}{\textit{Mon}}

Above, we investigated types indexed by abelian group expressions,
whose elements are taken from some group. We now switch from abelian
groups to commutative monoids.  Our main example is strings indexed by
``taint'' markers: when strings are appended their taint markers are
combined. Thanks to the information in the indexes, we are able to
prove an indefinability result (\thmref{thm:monoid-indefinability}).

We assume a single sort $\mathsf{M}$, with the monoid operations of
unit and multiplication, and the commutative monoid axioms. There is a
single primitive type $\tyPrimNm{M}$, with $\primTyArity(\tyPrimNm{M})
= [\mathsf{M}]$. For the index-erasure semantics, we assume some (not
necessarily commutative) monoid $M$, which we also write
multiplicatively. The collection of primitive operations
$\Gamma_{\mathit{Mon}}$ is simply the two monoid operations:
\begin{displaymath}
  \begin{array}{r@{\hspace{0.5em}:\hspace{0.5em}}l@{\hspace{2em}}r@{\hspace{0.5em}:\hspace{0.5em}}l}
    1 & \tyPrim{M}{1}, &
    (\cdot) & \forall a,b\mathord:\mathsf{M}.\ \tyPrim{M}{a} \to \tyPrim{M}{b} \to \tyPrim{M}{a b}
  \end{array}
\end{displaymath}
which we interpret using the monoid structure of $M$, giving an
environment $\eta_\Mon \in \ctxtSem{\Gamma_\Mon}$.

For the sets of relational environments, we make use of the
construction in \autoref{sec:constr-rel-env}, setting the chosen model
$\mathcal{M}$ to be the \emph{centre} $Z(M)$ of $M$ (recall that the
centre of a monoid is the set of all elements that commute with every
other element). We then set $R_{\tyPrimNm{M}}(m) = \{ (n,mn) \sepbar n
\in M \}$ and $R^\bullet_{\tyPrimNm{M}} = \emptyset$. By setting the
model $\mathcal{M}$ to only be the elements of $M$ that commute with
everything else, we ensure that the following lemma holds:
\begin{lemma}
  For all $\Delta$ and $\rho \in \relEnv{E}_\Mon(\Delta)$,
  $(\eta_\Mon,\eta_\Mon) \in \rsem{\Gamma_\Mon}{\relEnv{E}}\rho$.
\end{lemma}

\begin{example}
  By selecting a suitable monoid $M$, we can simulate a programming
  language that annotates strings with taint markers. Let $\Sigma$ be
  a set of characters from some alphabet, and let $T$ be a set of
  taint markers. Set $M = \Sigma^* \times
  \mathcal{MS}_{\mathit{fin}}(T)$, where $\mathcal{MS}_{\mathit{fin}}$
  denotes the finite power set operator. Since both $\Sigma^*$ and
  $\mathcal{MS}_{\mathit{fin}}(T)$ are monoids (the latter with
  multiset union as the monoid multiplication, which we write as
  $\uplus$), $M$ is itself a monoid. The centre of $M$, $Z(M)$,
  consists of all elements of the form $([], t)$.

  As an example of a free theorem for this instantiation of our
  general framework, consider programs of the following type:
  \begin{displaymath}
    ; \Gamma_\Mon \vdash N : \forall a,b\mathord:\mathsf{M}.\ \tyPrim{M}{a} \to \tyPrim{M}{b} \to \tyPrim{M}{aa}
  \end{displaymath}
  From this type we can derive the following free theorem about
  $N$. For all $a, b \in \mathcal{MS}_{\mathit{fin}}(T)$, $(s_1,t_1),
  (s_2,t_2) \in \Sigma^*$,
  \begin{displaymath}
    \begin{array}{l}
      \pi_2(\tmSem{N}\ (s_1,t_1 \uplus a)\ (s_2,t_2 \uplus b)) = \\
      \hspace{2em}\pi_2(\tmSem{N}\ (s_1,t_1)\ (s_2,t_2)) \uplus a \uplus a
    \end{array}
  \end{displaymath}
  If we taint the first input by $a$ and the second by $b$, then two
  copies of the taint $a$ will flow to the output, and none of taint
  $b$.
\end{example}

% if $b$ does not appear in the output, then it cannot have been used in
% the program. Turn this into a theorem...

% If forall i1...,in. M<e0> \to ... \to M<em> -> M<e>
% and e0 is not a factor of e, then we can get rid of the first input

% It ought to be invariant under its first input... Set R^\bullet to
% be the everywhere relation....

% prove this by using a relational environment that excludes

% \begin{theorem}
%   \begin{equation}
%     \label{eq:fo-monoid-type-insensitive}
%     ; \Gamma_\Mon \vdash M : \forall i_1,\dots,i_m\mathord:\mathsf{M}.\ \tyPrim{M}{e_1} \to \dots \to \tyPrim{M}{e_n} \to \tyPrim{M}{e}
%   \end{equation}
%   If all solutions to the matrix equation below have a row of zeros at
%   position $i$, then $\tmSem{M}$ is insensitive to 
% \end{theorem}

% as long as $e_1$ is not derivable from the other indexes, and the
% output one is, then the program is insensitive to its first argument:
% if $e_1 \not\in \Free(\{e_2,...,e_n\})$ and $e \in
% \Free(\{e_2,...,e_n\})$, then $M$ is insensitive to its first
% argument.

% \begin{proof}
%   Use the relational environment generated by the set $\{e_2, ...,
%   e_n\}$. Then the relational interpretation of the first argument's
%   type will be the total relation, and the remainder will be in the
%   identity relation. So for all $x_1,x_1' \in M$, and $x_2,...,x_n \in
%   M$ we have $\tmSem{M}\eta_\Mon\ x_1\ x_2...x_n = \tmSem{M}\eta_\Mon\
%   x'_1\ x_2...x_n$.
% \end{proof}

% Doesn't apply if we have:
% \begin{displaymath}
%   \forall a,b.\ \tyPrim{M}{a} \to \tyPrim{M}{a} \to \tyPrim{M}{b}
% \end{displaymath}
% but it ought to work... probably ought to state it in terms of minimal
% set of generators of the input: if there is a set of $e_i$ that
% generates the result type's index, then each of the inputs that it
% doesn't generate are not used by the function. Need the restrict to
% ones it doesn't generate because of the following:
% \begin{displaymath}
%   \forall a.\ \tyPrim{M}{a} \to \tyPrim{M}{a} \to \tyPrim{M}{a}
% \end{displaymath}
% we don't know which of the inputs will be used...

% Problem: what about
% \begin{displaymath}
%   \forall a.\ \tyPrim{M}{a} \to \tyPrim{M}{a^2} \to \tyPrim{M}{a^2}
% \end{displaymath}
% The set $\{a^2\}$ generates the output, and doesn't generate $a$, but
% a program of this type could easily use either of the inputs...

% Relational environment is:
% \begin{displaymath}
%   \rho\ \tyPrimNm{M}\ (e) = \left\{
%     \begin{array}{ll}
%       \id_M & \textrm{if }e = a^2 \\
%       M \times M & \textrm{otherwise}
%     \end{array}
%   \right.
% \end{displaymath}

% In general, we are just setting $R^\bullet_{\tyPrimNm{M}} = M \times
% M$. However, this is not supported by the monoid multiplication
% operation: if the two inputs are not in generated set, the output
% still might be, and is expected to be equal... So we need some notion
% of \emph{not} being generated by the set... free structures that avoid
% some argument?

% If we restrict to just variables, then we might be OK... otherwise,
% I'm stuck.

% The multiplication operation does not satisfy this spec. The
% generating set needs to be closed under decomposition....

% so we need to say that, somehow, the redundant arguments' indexes
% never contribute anything to the result type's index.

% Maybe: if an $e_i$ is not present in any of the minimal sets of
% generators that produces the output type, then it is redundant.

% Let $S = \{ G \subseteq \{e_1,...,e_n\} \sepbar e \in \Free(G) \}$.

% $S' = $ minimal elements of $S$: all the elements of $S$ that do not
% have anything beneath them in the subset ordering. Then if some $e_j$
% is not present in any of these,

\begin{theorem}
  \label{thm:monoid-indefinability}
  Let $m$ and $n$ be natural numbers. For all $n$-by-$m$-matricies of
  natural numbers $A$ and $m$-by-$1$ matricies of natural numbers $b$,
  there exists a program:
  \begin{equation}
    ; \Gamma_\Mon \vdash M : \forall i_1,\dots,i_m\mathord:\mathsf{M}.\ \tyPrim{M}{e_1} \to \dots \to \tyPrim{M}{e_n} \to \tyPrim{M}{e}
  \end{equation}
  where $e_j = i_1^{A_{j1}}\dots i_m^{A_{jm}}$ and $e = i_1^{b_1}\dots
  i_m^{b_m}$, iff there is a solution $\vec{x} = (x_1,...,x_n)$ in the
  natural numbers to $A x^\top = b$.
\end{theorem}

\begin{example}
  By \thmref{thm:monoid-indefinability}, it is impossible to write a
  program with the type $\forall m_1,m_2 \mathord: \mathsf{M}.\
  \tyPrim{M}{m_1m_2} \to \tyPrim{M}{m_1}$, showing that it is
  impossible to convert a doubly tainted value to a singly tainted
  one.

  In contrast to the case of abelian group indexed types, the lack of
  inverses in the monoid indexed setting means that the type $\forall
  m_1,m_2\mathord:\mathsf{M}.\ \tyPrim{M}{m_1m_2} \to \tyPrim{M}{m_2}
  \to \tyPrim{M}{m_1}$ is also uninhabited. There are no naturals
  $a_1,a_2$ such that $(m_1m_2)^{a_1}m_2^{a_2} \equiv m_1$.
\end{example}

% \fixme{Finish this} We can further assume that there are other
% operations that distribute over the multiplication of the monoid, for
% example semi-ring operations (e.g.~relational algebra).

%%% Local Variables:
%%% TeX-master: "paper"
%%% End: