### 2012-07-06 paper/section-singleton.tex

\section{Singleton Types}
\label{sec:singleton-types}

\newcommand{\Sing}{\mathit{Sing}}

\begin{enumerate}
\item Introduce the general idea of singleton types as a way of
  reifying information held in the indexes into the ``programming''
  world.
\item Helps maintain a separation between run-time and compile time
  data.
\end{enumerate}

Singleton types differ from the ``variation'' types that we have
considered so far because they nail down exactly the underlying value
instead of describing how it varies.

Indexing theory contains all the integers as constants, along with
$+$, $-$ and $*$. Normal arithmetic laws for these
operations. Program-level operations:
\begin{displaymath}
  \begin{array}{@{}c@{\hspace{0.5em}:\hspace{0.5em}}l}
    \underline{z} & \tyPrim{int}{z} \\ % FIXME: only need zero and one
    (+) & \forall z_1,z_2\mathord:\mathsf{integer}.\ \tyPrim{int}{z_1} \to \tyPrim{int}{z_2} \to \tyPrim{int}{z_1 + z_2} \\
    (-) & \forall z_1,z_2\mathord:\mathsf{integer}.\ \tyPrim{int}{z_1} \to \tyPrim{int}{z_2} \to \tyPrim{int}{z_1 - z_2} \\
    (*) & \forall z_1,z_2\mathord:\mathsf{integer}.\ \tyPrim{int}{z_1} \to \tyPrim{int}{z_2} \to \tyPrim{int}{z_1 * z_2}
  \end{array}
\end{displaymath}

\begin{enumerate}
\item Given an algebraic theory, and a model, we can define a
  singleton type system.
\item Need to carefully define this in general, and give the
  appropriate relational environments
\item Give the integers (freely generated monoid) as an example
\end{enumerate}

\begin{lemma}
  For all $\Delta$ and $\rho \in \relEnv{E}_\Sing(\Delta)$, $(\eta_\Sing, \eta_\Sing) \in \rsem{\Gamma}{\relEnv{E}_\Sing}{\rho}$.
\end{lemma}

Due to the precision of the index expressions, we obtain a very
precise characterisation of the definable programs under the singleton
typing discipline:
\begin{theorem}
  There exists a program
  \begin{displaymath}
    -; \Gamma_\Sing \vdash M : \forall i_1,...,i_m\mathord:\mathsf{S}.\ \tyPrim{S}{e_1} \to ... \to \tyPrim{S}{e_n} \to \tyPrim{S}{e}
  \end{displaymath}
  if and only if there exists an index expression
  $i'_1\mathord:\mathsf{S}, ..., i'_n\mathord:\mathsf{S} \vdash e' :
  \mathsf{S}$ such that $(e_1,...,e_n)^*e' \equiv e$.
\end{theorem}

\begin{proof}
  (If) Use the index expression $e'$ to define $M$, replacing each
  index operation with its program-level counterpart.

  (Only if) \fixme{use a special relational environment...}
\end{proof}

% \subsection{Predicate Types}
% \label{sec:predicate-types}

% \begin{enumerate}
% \item Let the indexing theory be predicates on the underlying value
% \item With the laws of boolean algebra
% \item Singleton types are a special case (i.e. with equality to a fixed value)
% \item But also have $\land$, $\lor$, $\lnot$ (and derived $\Rightarrow$)
% \item A pair of things is related if they both satisfy the predicate
%   (and are equal?)
% \item Need a special type for comparisons?
%   \begin{displaymath}
%     (<) : \forall z_1,z_2:\mathsf{integer}.\ \tyPrim{int}{z_1} \to \tyPrim{int}{z_2} \to (\tyPrim{int}{< z_1} \times 
%   \end{displaymath}
% \end{enumerate}

%%% Local Variables:
%%% TeX-master: "paper"
%%% End: