### 2012-07-06 paper/section-info-flow.tex
Updates to the logical information flow section

\section{Logical Information Flow}
\label{sec:information-flow}

\newcommand{\infoflow}{\mathit{Log}}

We now apply our general framework to types that are indexed by
logical propositions. By including a primitive type that represents
logical truth, we can recover--through a construction due to Tse and
Zdancewic \cite{tse04translating}--strong information flow properties
of programs. As a result of our general framework being parameterised
by the choice of equational theory, we can alter the logic that we use
for reasoning about type equality, and hence alter the information
flow properties of the system.

We first recall the concept of information flow. A function $f : A
\times B \to C$ is said to not allow information to flow from its
second argument to the output if for all $b,b' \in B$ and all $a \in
A$, $f(a,b) = f(a,b')$. If we think of the $B$ argument as
representing high-security information, then we have stated that $f$
does not allow the high-security input to flow to the low security
output. Information flow can be seen as a kind of invariance property
of programs, and so our relational interpretation of types is well
tailored to proving this kind of property.

As was thoroughly described by Sabelfeld and Sands
\cite{sabelfeld01per}, information flow can be captured semantically
through the use of partial equivalence relations (PERs). Abadi,
Banerjee, Heintze and Riecke \cite{abadi99core} built a \emph{Core
  Calculus for Dependency}, using a type system based around a
security level indexed monad $T_lA$, using PERs to prove the
information flow properties. Tse and Zdancewic \cite{tse04translating}
showed that it is possible to translate Abadi et al.'s calculus into
System F, translating the monadic type $T_lA$ to $\alpha_l \to A$ for
some free type variable $\alpha_l$, and making direct use of Reynolds'
abstraction theorem to prove information flow properties. For example,
if we use the type variable $\alpha_H$ to represent high-level
information, then the non-interference property of the function $f$
above could be expressed by the following System F type:
\begin{displaymath}
  A \to (\alpha_H \to B) \to C
\end{displaymath}
If a program cannot generate a value of type $\alpha_H$, then it
cannot access the value of $B$, and hence is insensitive to the actual
value. Relationships between security levels are captured by
postulating functions $\alpha_{l_1} \to \alpha_{l_2}$ whenever $l_1$
is a lower security level than $l_2$.

By making use of algebraically indexed types, we are able to refine
Tse and Zdancewic's translation by replacing the use of type variables
$\alpha_l$ with a primitive type $\tyPrim{T}{e}$ of representations of
the truth of a logical proposition $e$. The relationships between
security levels are now replaced by logical entailment, with the
consequence that we only have functions of type $\tyPrim{T}{e_1} \to
\tyPrim{T}{e_2}$ when $e_1$ entails $e_2$. Inspired by the PER
interpretation of information flow mentioned above, we instantiate our
general relational framework to interpret $\tyPrimNm{T}{e}$ as the
identity relation when $e$ is true, and the empty relation when $e$ is
false. In doing so, we can recover the information flow properties of
Abadi et al. and Tse and Zdancewic.

\paragraph{Instantiation of the General Framework}
For this instantiation of our framework, we assume a single indexing
sort $\mathsf{prop}$. For now, we assume that this sort supports
meet-semilattice structure: we have a binary operation $\land$ and a
constant $\top$, where $\land$ is commutative, associative and has
$\top$ as a unit, and $\land$ is idempotent: so $p \land p =
p$. Recall that we can define a partial order structure on any
meet-semilattice by setting $p \leq q$ when $p = p \land q$. We take
an arbitrary meet-semilattive $L$ for our chosen model.

% Below, we will consider additional structure on the sort
% $\mathsf{prop}$ and the model $L$

We assume a single primitive type $\tyPrimNm{T}$, with
$\primTyArity(\tyPrimNm{T}) = [\mathsf{prop}]$ and index-erasure
semantics $\tyPrimSem{\tyPrimNm{X}} = \{ * \}$. Thus values of type
$\tyPrimNm{T}{e}$ have no real run-time content; their meaning is
given by the relational semantics. For the relational environments
$\relEnv{E}_\infoflow$, we use the general construction in
\autoref{sec:constr-rel-env}, with our chosen model $L$ and the
following relational interpretations:
\begin{displaymath}
  \begin{array}{l@{\hspace{2em}}l}
    R^\bullet_{\tyPrimNm{T}} = \emptyset &
    R_{\tyPrimNm{T}}(e) = \{ (*,*) \sepbar e = \top \}
  \end{array}
\end{displaymath}
The primitive operations $\Gamma_\infoflow$ for this instantiation
reflect the meet-semilattice structure of the indexes:
\begin{displaymath}
  \begin{array}{l}
  \begin{array}{@{}c@{\hspace{2em}}c}
    \mathrm{truth} : \tyPrim{T}{\top} &
    \mathrm{and}   : \forall p, q\mathord:\mathsf{prop}.\ \tyPrim{T}{p} \to \tyPrim{T}{q} \to \tyPrim{T}{p \land q}
  \end{array} \\
  \mathrm{up} : \forall p, q\mathord:\mathsf{prop}.\ \tyPrim{T}{p \land q} \to \tyPrim{T}{p}
\end{array}
\end{displaymath}
The combination of the \TirName{TyEq} rule and the primitive
$\mathrm{up}$ operation allow for logical entailment to be reflected
in programs. Each of the primitive operations has a trivial
interpretation, due the index-erasure interpretation of
$\tyPrim{T}{e}$ as a one-element set, giving an environment
$\eta_\infoflow \in \ctxtSem{\Gamma}$. Less trivially, we have this
lemma:
\begin{lemma}\label{lem:environments-information-flow}
  For all $\Delta$ and $\rho \in \relEnv{E}_\infoflow$, $(\eta_\infoflow, \eta_\infoflow) \in \rsem{\Gamma_\infoflow}{\relEnv{E}_\infoflow}{\rho}$.
\end{lemma}

It is also possible to think of more complicated logics, as long as
they include

\begin{theorem}
  Inexpressibility result
\end{theorem}

\paragraph{Application to Information Flow}
We now think of logical expressions as ``composite principals''. That
is, propositional variables represent various principals that are
combined using the logical connectives. We interpret ``truth'' for
principals as stating that that principal is satisfied with the
current state of affairs, and that

We adapt Tse and
Zdancewic's translation of Abadi et al.'s monadic type to our
setting. We define a type synonym, for each type $A$ and expression
$e$ of sort $\mathsf{principal}$:
\begin{displaymath}
  T_eA = \tyPrim{T}{e} \to A
\end{displaymath}
For every principal $e$, we can endow the types $T_e-$ with the
structure of a monad. This is due to the fact that it is an instance
of the ``environment'' (or ``reader'') monad \cite{jones95functional}.

For any extension of the logical theory...

\begin{example}
  Different notions of entailment lead to different notions of
  non-interference.
\end{example}

\begin{theorem}
  Type non-inhabitation result
\end{theorem}

\begin{theorem}
  Type isomorphism theorem, making sure to emphasi
\end{theorem}

% In the previous two sections, we looked at algebraically-indexed types
% whose relational interpretations used functional relations, with the
% consequence that the derived free theorems express invariance
% properties.

% In this section we examine a type

% We now look at another application of algebraically indexed types that
% does not involve the kinds of functional relationships between values
% that we have looked at the previous two sections. Rather, we use the
% relational interpretation to restrict the elements of types

% \begin{enumerate}
% \item Make sure the logic is at least a meet semi-lattice. We have a
%   single indexing sort $\mathsf{prn}$. There is a binary operation
%   $\land$ which is associative and commutative and has unit $\top$,
%   and also the binary operation is idempotent: $p \land p = p$.
% \item Have a single primitive type $\tyPrimNm{T}$, with
%   $\primTyArity(\tyPrimNm{T}) = [\mathsf{prn}]$. The index-erasure
%   semantics of this type interprets it simply as 
% \end{enumerate}

% \begin{enumerate}
% \item Also makes use of the relational aspect, the previous stuff
%   could make do with unary logical predicates as the interpretations
%   of types, but the information flow properties make use of the
%   relational interpretation.
% \item Can also obtain a type non-inhabitation result based on logical
%   derivability
% \end{enumerate}

% \fixme{
%   \begin{enumerate}
%   \item Make the definition of control of information flow clearer
%   \item Make the example a bit more compelling
%   \item Put in some examples of composite principals formed from
%     boolean logic, and the use of boolean algebra to do reasoning in
%     the types. List the laws of boolean algebra, and the laws of
%     Heyting algebra. Need to explain the difference between 
%   \item Possibly use singleton types to be able to do more interesting
%     security policies?
%   \item Put in the formal statement of the information flow property
%     and the proof.
%   \end{enumerate}
% }

% An extreme example of invariance under change of representation is the
% tracking of information flow through programs. If a program does not
% depend upon a particular piece of information, then it will be
% invariant under \emph{all} changes of representation of this
% information. As thoroughly described by Sabelfeld and Sands
% \cite{sabelfeld01per}, information flow can be captured semantically
% through the use of partial equivalence relations (PERs). Interpreting
% types as PERs forms an instance of Reynolds' approach to abstraction
% and representation independence. Abadi, Banerjee, Heintze and Riecke
% \cite{abadi99core} built a \emph{Core Calculus for Dependency}, using
% a type system based around a security level indexed monad $T_lA$. Tse
% and Zdancewic \cite{tse04translating} showed that it is possible to
% translate Abadi et al.'s calculus into System F, translating the
% monadic type $T_lA$ to $\alpha_l \to A$ for some free type variable
% $\alpha_l$, and making direct use of Reynolds' abstraction theorem to
% prove information flow properties. We now show how a refined variant
% of Tse and Zdancewic's translation can be expressed via
% algebraically-indexed types.

% We assume a single indexing sort $\mathsf{principal}$, intended to
% represent (possibly composite) principals in a system. Principals are
% combined using the connectives of boolean logic ($\top$, $\bot$,
% $\lor$, $\land$, $\lnot$). For the equational theory of principals, we
% assume all the axioms of boolean algebra, and semantically we
% interpret composite principals as boolean values: either true ($\top$)
% or false ($\bot$). We introduce a single primitive type, indexed by
% expressions of sort $\mathsf{principal}$, $\tyPrim{T}{e}$, with the
% following interpretations:
% \begin{displaymath}
%   \begin{array}{l@{\hspace{0.5em}=\hspace{0.5em}}l}
%     \tySem{\tyPrim{T}{e}} & \{*\} \\
%     \rsem{\tyPrim{T}{e}}{}\rho & \left\{
%       \begin{array}{ll}
%         \{(*,*)\} & \textrm{if }\sem{e}\rho = \top \\
%         \{\}      & \textrm{if }\sem{e}\rho = \bot
%       \end{array}
%       \right.
%   \end{array}
% \end{displaymath}
% The relational interpretation of $\tyPrim{T}{e}$ forces this type to
% only be inhabited when the boolean expression $e$ evaluates to true
% under the environment $\rho$. To write programs,

% \begin{example}
%   Suppose we have a program with the following type:
%   \begin{displaymath}
%     \forall p,q\mathord:\mathsf{principal}. \tyPrim{T}{e} \to T_p(\tyPrimNm{unit} + \tyPrimNm{unit}) \to T_q(\tyPrimNm{unit} + \tyPrimNm{unit})
%   \end{displaymath}
%   where $e$ is some boolean algebra expression involving $p$ and
%   $q$. We will show in \autoref{sec:instantiations} that programs with
%   this type are non-constant (i.e.~depend on their input) if and only
%   if the formula $e \Rightarrow q \Rightarrow p$ is valid in boolean
%   logic.
% \end{example}

% \begin{enumerate}
% \item Treat the information flow thing as a type isomorphism, in the
%   same way that Tse and Zdancewic do
%   \begin{displaymath}
%     \begin{array}{l}
%     \forall \vec{p}\mathord:\mathsf{principal}.\ T_{e_1}\tyPrimNm{bool} \to T_{e_2}\tyPrimNm{bool}\\
%     \hspace{4em}\cong \\
%     \left\{
%       \begin{array}{ll}
%         \tyPrimNm{bool} \tyArr \tyPrimNm{bool} & \textrm{if }e_2 \vdash e_1 \\
%         \tyPrimNm{bool} & \textrm{if }e_2 \not\vdash e_1
%       \end{array}
%     \right.
%   \end{array}
%   \end{displaymath}
%   where $\tyPrimNm{bool} = \tyUnit + \tyUnit$.
% \item mention the issue with bogus tokens if we allow for a
%   call-by-name style of non-termination
% \item if we axiomatise intuitionistic logic instead of
%   boolean/classical logic, we get a different type isomorphism.
% \item The isomorphism relies upon $e_1 \vdash e_2 \Leftrightarrow e_1
%   \lor e_2 = e_1$.
% \item Relevant logic? Look at the Wikipedia page to see how to do
%   residuated (commutative) monoids on lattices with just
%   equations. With this, we will be able to do substructural logics
%   too. As long as we know that a particular axiomatisation is sound
%   and complete for some reasoning system, then we can state and prove
%   the non-interference isomorphism above.
% \item Compare to Tse and Zdancewic's addition of axioms to the logic.
% \item Internalising equality? The logical structure can then be used
%   on types as well as everything else. Need the special rules for
%   equality... (see the Fomega paper).
% \end{enumerate}

% Go a step further by considering equivalence relations on word32, and
% their lattice structure. Link this to quantative information
% flow. Will also need to include singleton types.

%%% Local Variables:
%%% TeX-master: "paper"
%%% End: