### 2012-07-09 paper/paper.tex

\documentclass[preprint]{sigplanconf}

\usepackage[usenames]{color}
\usepackage[numbers]{natbib}
\definecolor{citationcolour}{rgb}{0,0.4,0.2}
\definecolor{linkcolour}{rgb}{0,0,0.8}
\usepackage{hyperref}
\hypersetup{colorlinks=true,
            urlcolor=linkcolour,
            linkcolor=linkcolour,
            citecolor=citationcolour,
            pdftitle=Relational Parametricity for Algebraically-Indexed Types,
            pdfauthor={Robert Atkey, Neil Ghani, Patricia Johann, Andrew Kennedy},
            pdfkeywords={}}  
\def\sectionautorefname{Section}
\def\subsectionautorefname{Section}
\def\subsubsectionautorefname{Section}

\title{Abstraction and Invariance for Algebraically Indexed Types}

\authorinfo{Robert Atkey\and Patricia Johann}
           {University of Strathclyde}
           {\{Robert.Atkey,Patricia.Johann\}@strath.ac.uk}

\authorinfo{Andrew Kennedy}
           {Microsoft Research Cambridge}
           {akenn@microsoft.com}

\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage[all]{xy}

\newcommand{\todo}[1]{[\textbf{TODO}: #1]}

\newcommand{\GL}[1]{\mathrm{GL}_#1}
\newcommand{\SynGL}[1]{\mathsf{GL}_#1}
\newcommand{\SE}[1]{\mathsf{SE}_#1}
\newcommand{\SynSE}[1]{\mathsf{SE}_#1}
\newcommand{\Orth}[1]{\mathrm{O}_#1}
\newcommand{\SynOrth}[1]{\mathsf{O}_#1}
\newcommand{\Transl}[1]{\mathrm{T}_#1}
\newcommand{\SynTransl}[1]{\mathsf{T}_#1}
\newcommand{\Scal}{\mathrm{Scal}}
\newcommand{\SynScal}{\mathsf{Scal}}

% \newcommand{\GLtwo}{\mathrm{GL}_2}
% \newcommand{\SynGLtwo}{\mathsf{GL}_2}
% \newcommand{\GLone}{\mathrm{GL}_1}
% \newcommand{\SynGLone}{\mathsf{GL}_1}

\newcommand{\sepbar}{\mathrel|}

\newcommand{\Rel}{\mathrm{Rel}}
\newcommand{\relArrow}{\mathrel{\widehat\to}}
\newcommand{\relTimes}{\mathrel{\widehat\times}}
\newcommand{\relSum}{\mathrel{\widehat+}}
\newcommand{\Eq}{\mathrm{Eq}}

\newcommand{\setOfIntegers}{\mathbb{Z}}
\newcommand{\setOfBooleans}{\{\tmTT, \tmFF\}}

\newcommand{\id}{\mathrm{id}}

\newcommand{\SortSet}{\mathit{Sort}}
\newcommand{\IndexOpSet}{\mathit{IndexOp}}
\newcommand{\PrimTypeSet}{\mathit{PrimType}}
\newcommand{\IndexAxiomSet}{\mathit{IndexAx}}

\newcommand{\indexOp}[1]{\texttt{#1}}
\newcommand{\idxTms}[2]{\mathrm{IdxTm}(#1 \vdash #2)}
\newcommand{\idxTmAS}[1]{\mathrm{IdxTm}(#1)}

\newcommand{\tyInt}{\texttt{int}}
\newcommand{\tyBool}{\texttt{bool}}
\newcommand{\tyUnit}{\texttt{unit}}
\newcommand{\tyReal}{\texttt{real}}
\newcommand{\tyPrim}[2]{\textup{\texttt{#1}}\langle #2 \rangle}
\newcommand{\tyPrimNm}[1]{\textup{\texttt{#1}}}
\newcommand{\primTyArity}{\mathrm{tyArity}}
\newcommand{\indexOpArity}{\mathrm{opArity}}

\newcommand{\tyArr}{\to}
\newcommand{\tyProduct}{\times}
\newcommand{\tyX}[1]{\texttt{X}\langle #1 \rangle}
\newcommand{\isType}{\textsf{ type}}
\newcommand{\isCtxt}{\textsf{ ctxt}}
\newcommand{\Ty}{\textsf{Ty}}
\newcommand{\ty}{\textsf{ty}}

\newcommand{\tmTT}{\texttt{tt}}
\newcommand{\tmFF}{\texttt{ff}}

\newcommand{\relEnv}[1]{\mathcal{#1}}
\newcommand{\tySem}[1]{\llbracket #1 \rrbracket^{\mathcal{T}}}
\newcommand{\ctxtSem}[1]{\llbracket #1 \rrbracket^{\mathcal{C}}}
\newcommand{\tmSem}[1]{\llbracket #1 \rrbracket^{\mathit{tm}}}
\newcommand{\tyPrimSem}[1]{\llbracket #1 \rrbracket^{\mathcal{T}_0}}
\newcommand{\rsem}[3]{\llbracket #1 \rrbracket^{\mathcal{R}}_{#2}{#3}}
\newcommand{\extends}[2]{\mathsf{ext}(#1,#2)}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newcommand{\lemref}[1]{\hyperref[#1]{Lemma \ref*{#1}}}
\newcommand{\thmref}[1]{\hyperref[#1]{Theorem \ref*{#1}}}
\newcommand{\defref}[1]{\hyperref[#1]{Definition \ref*{#1}}}
\newcommand{\propref}[1]{\hyperref[#1]{Proposition \ref*{#1}}}
\newcommand{\corref}[1]{\hyperref[#1]{Corollary \ref*{#1}}}
\newcommand{\conref}[1]{\hyperref[#1]{Conjecture \ref*{#1}}}
\newcommand{\exref}[1]{\hyperref[#1]{Example \ref*{#1}}}
\newcommand{\statementref}[1]{\hyperref[#1]{Statement \ref*{#1}}}

\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\isDefinedAs}{\stackrel{\mathit{def}}=}

\newcommand{\Geom}{\mathit{Geom}}

\newtheoremstyle{examplestyle}
  {\topsep}  % space above
  {\topsep}  % space below
  {\normalfont}% name of font to use in the body of the theorem
  {0em}% measure of space to indent
  {\bfseries}% name of head font
  {.}% punctuation between head and body
  {5pt plus 1pt minus 1pt}% space after theorem head; " " = normal interword space
  {}% Manually specify head
\theoremstyle{examplestyle}
\newtheorem{example}{Example}

\newtheoremstyle{restatementstyle}
  {\topsep}  % space above
  {\topsep}  % space below
  {\itshape}% name of font to use in the body of the theorem
  {0em}% measure of space to indent
  {\bfseries}% name of head font
  {.}% punctuation between head and body
  {5pt plus 1pt minus 1pt}% space after theorem head; " " = normal interword space
  {\thmname{#1} \thmnote{(Restatement of #3)}}% Manually specify head
\theoremstyle{restatementstyle}
\newtheorem{restateLemma}{Lemma}
\newtheorem{restateTheorem}{Theorem}

\newcommand{\fixme}[1]{\textbf{FIXME: #1}}

\begin{document}

\maketitle

\begin{abstract}
  Reynolds' relational parametricity provides a powerful way to reason
  about programs in terms of invariance under changes of data
  representation. A dazzling array of applications of Reynolds' theory
  exists, exploiting invariance to yield ``free theorems'',
  uninhabitation results, and encodings of algebraic datatypes.
  Outside computer science, invariance is a common theme running
  through many areas mathematics and physics. For example, the area of
  a triangle is unaltered by rotation or flipping. If we scale a
  triangle, then we scale its area, maintaining an invariant
  relationship between the two. The transformations under which
  properties are invariant are often organised into groups, with the
  algebraic structure reflecting the composability and invertibility
  of transformations.

  In this paper, we investigate programming languages whose types are
  indexed by algebraic structures such as groups of geometric
  transformations. Other examples include types indexed by
  principals--for information flow security--and types indexed by
  distances--for analysis of analytic uniform continuity
  properties. Following Reynolds, we prove a general abstraction
  theorem that covers all these instances. Consequences of our
  abstraction theorem include free theorems expressing invariance
  properties of programs, type isomorphisms based on invariance
  properties, and non-definability results indicating when certain
  algebraically indexed types are uninhabited or only inhabited by
  trivial programs.  Our results have been fully formalised in Coq.
\end{abstract}

\category{D.1.1}{Programming techniques}{Applicative (functional)
  programming} \category{D.2.4}{Software Engineering}{Software/Program
  Verification} \category{D.3.3}{Programming Languages}{Language
  Constructs and Features---Data types and structures}

\terms
  Languages, Theory, Types

\keywords
  parametricity, units of measure, dimensional analysis, invariance, computational geometry

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{section-introduction.tex}
\input{geometry-twocolumn.tex}
\input{section-general-framework.tex}
\input{section-instantiations.tex}
\input{section-info-flow.tex}
\input{section-monoid.tex}
\input{section-singleton.tex}
\input{section-metric.tex}
\input{section-discussion.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{abbrvnat}
\bibliography{paper}

\newpage
\appendix

\input{appendix-proofs.tex}

\end{document}