### 2012-07-09 paper/section-info-flow.tex

\section{Logical Information Flow}
\label{sec:information-flow}

\newcommand{\infoflow}{\mathit{Log}}

We now apply our general framework to types that are indexed by
logical propositions. By including a primitive type that represents
logical truth, we can recover--through a construction due to Tse and
Zdancewic \cite{tse04translating}--strong information flow properties
of programs. As a result of our general framework being parameterised
by the choice of equational theory, we can alter the logic that we use
for reasoning about type equality, and hence alter the information
flow properties of the system.

We first recall the concept of information flow. A function $f : A
\times B \to C$ is said to not allow information to flow from its
second argument to the output if for all $b,b' \in B$ and all $a \in
A$, $f(a,b) = f(a,b')$. If we think of the $B$ argument as
representing high-security information, then we have stated that $f$
does not allow the high-security input to flow to the low security
output. Information flow can be seen as a kind of invariance property
of programs, and so our relational interpretation of types is well
tailored to proving this kind of property.

As described by Sabelfeld and Sands \cite{sabelfeld01per}, information
flow can be captured semantically by partial equivalence relations
(PERs). Abadi, Banerjee, Heintze and Riecke \cite{abadi99core} built a
\emph{Core Calculus for Dependency}, using a type system based around
a security level indexed monad $T_lA$, using PERs to prove the
information flow properties. Tse and Zdancewic \cite{tse04translating}
translated Abadi et al.'s calculus into System F, translating the
monadic type $T_lA$ to $\alpha_l \to A$ for some free type variable
$\alpha_l$, and using Reynolds' abstraction theorem to prove
information flow properties. For example, if the type variable
$\alpha_H$ represents high-level information, then the
non-interference property of the function $f$ could be expressed by
the System F type $A \to (\alpha_H \to B) \to C$. If a program cannot
generate a value of type $\alpha_H$, then it cannot access the value
of $B$, and hence is insensitive to the actual value. Relationships
between security levels are captured by postulating functions
$\alpha_{l_1} \to \alpha_{l_2}$ whenever $l_1$ is a lower security
level than $l_2$.

Using algebraically indexed types, we refine Tse and Zdancewic's
translation by replacing each type variable $\alpha_l$ with a
primitive type $\tyPrim{T}{e}$ of representations of the truth of a
logical proposition $e$. The relationships between security levels are
now replaced by logical entailment, so we only have functions of type
$\tyPrim{T}{\phi} \to \tyPrim{T}{\psi}$ when $\phi$ entails
$\psi$. Inspired by the PER interpretation of information flow, we
instantiate our general relational framework to interpret
$\tyPrimNm{T}{e}$ as the identity relation when $e$ is true, and the
empty relation when $e$ is false. In doing so, we can recover the
information flow properties of Abadi et al. and Tse and Zdancewic.

\paragraph{Instantiation of the General Framework}
We assume a single indexing sort $\mathsf{prop}$ and assume the
operations and equations of boolean algebra. Thus we have constants
$\top$, $\bot$ and binary operators $\land$, $\lor$ with the axioms of
a bounded lattice, and a unary complementation operator $\lnot$. We
will use $\phi, \psi$ to stand for index expressions of sort
$\mathsf{prop}$. We use an equational presentation to fit with our
general framework, but note that we can define an order on index
expressions as $\phi \leq \psi$ when $\phi = \phi \land \psi$.

We have a single primitive type $\tyPrimNm{T}$, with
$\primTyArity(\tyPrimNm{T}) = [\mathsf{prop}]$ and index-erasure
semantics $\tyPrimSem{\tyPrimNm{X}} = \{ * \}$. Thus values of type
$\tyPrim{T}{\phi}$ have no run-time content; their only meaning is
given by the relational semantics. For the relational environments
$\relEnv{E}_\infoflow$, we use the compositional construction in
\autoref{sec:constr-rel-env}, with an arbitrary boolean algebra $L$
and $R_{\tyPrimNm{T}}(x) = \{ (*,*) \sepbar x = \top \}$.  The
primitive operations $\Gamma_\infoflow$ reflect logical consequence:
\begin{displaymath}
  \begin{array}{l}
  \begin{array}{@{}c@{\hspace{2em}}c}
    \mathrm{truth} : \tyPrim{T}{\top} &
    \mathrm{and}   : \forall p, q\mathord:\mathsf{prop}.\ \tyPrim{T}{p} \to \tyPrim{T}{q} \to \tyPrim{T}{p \land q}
  \end{array} \\
  \mathrm{up} : \forall p, q\mathord:\mathsf{prop}.\ \tyPrim{T}{p \land q} \to \tyPrim{T}{p}
\end{array}
\end{displaymath}
The combination of the \TirName{TyEq} rule and the primitive
$\mathrm{up}$ operation allow for logical entailment to be reflected
in programs: if we have $\phi \leq \psi$ and $M : \tyPrim{T}{\phi}$
then $\mathrm{up}\ M : \tyPrim{T}{\psi}$. Each of the primitive
operations has a trivial interpretation, due the index-erasure
interpretation of $\tyPrim{T}{\phi}$ as a one-element set, giving an
environment $\eta_\infoflow \in \ctxtSem{\Gamma}$. Less trivially, we
have this lemma:
\begin{lemma}\label{lem:environments-information-flow}
  For all $\Delta$ and $\rho \in \relEnv{E}_\infoflow$, $(\eta_\infoflow, \eta_\infoflow) \in \rsem{\Gamma_\infoflow}{\relEnv{E}_\infoflow}{\rho}$.
\end{lemma}

\paragraph{Information Flow} We think of logical expressions as
``composite principals''. That is, propositional variables represent
atomic principals that are combined using the logical connectives. We
interpret ``truth'' for principals as stating that a principal is true
when satisfied with the current state of affairs. Thus a relationship
$\phi \leq \psi$ indicates that satisfaction of the composite
principal $\phi$ implies satisfaction of the composite principal
$\psi$. In terms of security levels, the ordering is reversed: if a
high security principal is satisfied, then all of their underlings
must also be satisfied.

To study information flow, we adapt Tse and Zdancewic's translation of
Abadi et al.'s monadic type to our setting. We define a type
abbreviation $T_\phi A = \tyPrim{T}{\phi} \to A$, where $A$ is a type
and $\phi$ is an expression of sort $\mathsf{prop}$. For every $\phi$,
we can endow the types $T_\phi-$ with the structure of a monad. This
is due to the fact that it is an instance of the ``environment'' (or
``reader'') monad \cite{jones95functional}. We read the types
$T_\phi A$ as data of type $A$ ``protected'' by the principal
$\phi$.

As a consequence of the relational interpretation given above, it
follows that if we have an index expression $\phi$ that is interpreted
as some value other than $\top$ in a relational environment $\rho$,
then for all $x,x' \in \tySem{T_\phi\tyPrimNm{bool}}$, we have $(x,x')
\in \rsem{T_\phi\tyPrimNm{bool}}{\relEnv{E}_\infoflow}{\rho}$. In
terms of the reading above, if a principal is dissatisfied
(i.e.,~$\phi \not= \top$), then data protected by this principal is
indistinguishable from any other data. Thus a program cannot get
access to the exact value. From this and \thmref{thm:abstraction}, we
obtain the following information flow result:

\begin{theorem}
  Let $\phi$ and $\psi$ be index expressions of sort $\mathsf{prop}$
  in some indexing context $\Delta$, such that $\psi \not\leq
  \phi$. Then for all programs
  \begin{displaymath}
    \Delta; \Gamma_\infoflow,\Gamma \vdash M : T_\phi\tyPrimNm{bool} \to T_\psi\tyPrimNm{bool}
  \end{displaymath}
  and all terms $N_1,N_2$ of type $T_\phi\tyPrimNm{bool}$, $M\ N_1
  \stackrel{ctx}\approx M\ N_2$.  Thus there is no information flow
  from $M$'s input to its output.
\end{theorem}
This result also holds if we move to logics other than boolean
logic. For example, if our equational theory models intuitionistic
logic by taking the axioms of Heyting algebras, then the same non-flow
property for programs of type $T_{p\lor\lnot p}\tyPrimNm{bool} \to
T_\top\tyPrimNm{bool}$ holds, due to the lack of excluded middle. If
we take linear logic, then programs of type $T_{p\otimes
  p}\tyPrimNm{bool} \to T_p\tyPrimNm{bool}$ have no information flow
from their input, due to non-provability of $p \vdash p \otimes p$.

%%% Local Variables:
%%% TeX-master: "paper"
%%% End: