### 2012-07-10 paper/section-discussion.tex
Patty's changes to shorten the text

\section{Discussion}
\label{sec:discussion}

We presented a general framework for algebraically indexed types
and instantiated it to yield novel type systems for geometry, taint
tracking, logical information flow and distance-indexed types. Our
framework further demonstrates the power of relational reasoning about
typed programs. From \thmref{thm:abstraction}, we derived
interesting free theorems, type isomorphisms and non-definability
results. We conclude with some observations and suggestions for
further work.

\paragraph{Further Applications and Extensions} We have covered
several applications of algebraically indexed types in this paper, but
there are undoubtedly many more. Geometry for dimensions greater than
two is an obvious candidate, as are systems that are invariant
under different geometric groups (e.g.,~the Poincar\'{e} group for
relativity). Mathematical Physics is particularly rich in theories
that have some notion of invariance, and it will be exciting to pin
down the precise connections between these and type systems for which
%analogues of Reynolds' 
an abstraction theorem holds. Geometric theorem
proving is another %possible 
application. Harrison
\cite{harrison09without} comments on the pervasiveness of % usefulness of
invariance properties in this area. Programs in our framework
automatically satisfy invariance properties, removing the need for
{\em ad hoc} proofs. % of these facts.

In \autoref{sec:metric-types} we presented a type system with
distance-indexed types, and noted the similarity with Reed and
Pierce's system for non-expansivity \cite{reed10distance}. Exploring
the relationship--particularly with respect to probability
monads--also looks promising.

Type and effect analyses use types indexed by effect annotations with
algebraic structure (e.g.,~sets of read/write effect labels with an
idempotent monoid structure). Benton {\em et al.}~ have used relational
interpretations to prove effect-dependent equivalences
\cite{benton06reading}. An extension of our framework with
type-indexed types should be able to express their effect-indexed monads
and prove their equivalences.

Extending our framework with type dependency would also allow for
further applications. For example, we could consider a type of lists
of length $n$, indexed by elements of the permutation group
$S_n$. Bernardy et al.~have presented a general framework for
relational reasoning and an abstraction theorem for dependent types
\cite{bernardy12proofs}. However, they work with pure type systems,
which define type equality via untyped rewriting, so it is not
immediately obvious how to integrate arbitrary equational theories
into their framework.

\paragraph{Extension of the non-definability results} Our
non-definabilty results are currently restricted to the quantifier
free fragment of our framework, as defined in the statement of
\thmref{thm:abstraction}. This is a consequence of the
non-compositionality of the relational environments that we use to
prove non-definabilty results. Lifting this restriction is a key item
of future work.

\paragraph{Semantic Equality} In general, the semantic equality
$\stackrel{sem}\sim$ in \defref{def:semantic-equality} is not an
equivalence relation. If we assume that the interpretations of all the
primitive types are partial equivalence relations then semantic
equality is indeed an equivalence relation. However, this excludes the
geometry and distance-indexed examples. More generally, we can
consider relational interpretations that are \emph{difunctional}. (A
relation is difunctional if whenever $(x,y)$, $(x',y')$ and $(x,y')$
are in the relation then so is $(x',y)$.) Difunctionality is weaker
than being a PER, but still suffices to prove that semantic equality
is an equivalence relation. Hofmann \cite{hofmann08correctness} has
used difunctional relations in the setting of effect
analyses. Difunctionality covers all our examples except
distance-indexed types. Note that for both PERs and difunctional
relations we need to close the relational interpretation of
existential types under the appropriate property to ensure that all
types are interpreted as PERs/difunctional relations. For
distance-indexed types it is possible that a new notion of
equivalence, based on closeness, is required.

%%% Local Variables:
%%% TeX-master: "paper"
%%% End: