### 2012-10-24 paper/section-general-nondef.tex

\section{Non-definability for the General Framework}
\label{sec:general-nondef}

The non-definability results in the previous section required the use
of specially crafted models and relational interpretations. It is
reasonable to ask whether there is a general method for constructing
suitable models and relational interpretations to prove
non-definability results. In this section, we show that this is
possible for a large class of first-order types.

In \exref{ex:sqrt-root-nondef}, we showed that the type $\forall s
\mathord: \SynGL{1}.~\tyPrim{real}{s^2} \to \tyPrim{real}{s}$ only has
trivial inhabitants. Intuitively, this is because the index of the
result type ($s$) cannot be obtained from the index of the input type
($s^2$) using the abelian group operations and axioms. This
observation can be used to give a sufficient condition for
non-inhabitation for any instantiation of our general framework for
types of the form:
\begin{displaymath}
  \forall i_1\mathord:s_1,\ldots,i_m\mathord:s_m.~\tyPrim{X}{e_1} \to \ldots \to \tyPrim{X}{e_n} \to \tyPrim{X}{e}
\end{displaymath}
Roughly speaking, if this type is inhabited, then it must be the case
that the index expression $e$ can be generated from the set of index
expressions $\{e_1,...,e_n\}$.

We assume that we are working with an instantiation of the general
framework from \autoref{sec:a-general-framework} with a closed typing
context $\Gamma_{\mathit{ops}}$ describing the types of the primitive
operations and a chosen index-erasure semantics. We will use a special
relational interpretation built from the syntax of the indexing
expressions to prove our general non-definability result. For
simplicity, we assume that there is only one primitive type,
$\tyPrimNm{X}$, but the technique we describe here extends to the
general case. We also assume that $\tySem{\tyPrimNm{X}}$ is non-empty.

To state our general non-definability condition, we need to define the
set of index expressions generated by some finite set of index
expressions. Given a set $S$ of index expressions that are all
well-indexed in some index context $\Delta$, we define
$\mathrm{Gen}_0(S)$ to be the set of expressions that are built from
the elements of $S$ and the primitive index operations. To account for
the equations between indexing terms, we close under index expression
equivalence to get the set $\Gen(S) = \{ e \sepbar \exists s.\ \exists
e' \in \mathrm{Gen}_0(S).\ \Delta \vdash e \equiv e' : s \}$. Now a
type is \emph{well-generated} if it is closed and of the form:
\begin{displaymath}
  \forall i_1\mathord:s_1,\ldots,i_m\mathord:s_m.~\tyPrim{X}{e_1} \to \ldots \to \tyPrim{X}{e_n} \to \tyPrim{X}{e}
\end{displaymath}
and $e \in \Gen(\{e_1,...,e_n\})$.
\begin{theorem}\label{thm:general-nondef}
  Assume that the members of $\Gamma_{\mathit{ops}}$ are all of
  well-generated types. If there exists an $M$ with typing:
  \begin{displaymath}
    \Gamma_{\mathit{ops}} \vdash M : \forall i_1\mathord:s_1,...,i_m\mathord:s_m.~\tyPrim{X}{e_1} \to ... \to \tyPrim{X}{e_n} \to \tyPrim{X}{e}
  \end{displaymath}
  then $e \in \Gen(\{e_1,...,e_n\})$.
\end{theorem}
This theorem is usually applied in the contrapositive: if $e \not\in
\Gen(\{e_1,...,e_n\})$ then no such $M$ can exist.
\begin{proof}
  Let $\Delta = i_1\mathord:s_1,...,i_m\mathord:s_m$ be the index
  context constructed from the universally quantified type variables
  in the type of $M$.  To interpret the index expressions, we use the
  free model over the variables in $\Delta$ constructed from the
  syntax. This model assigns to each sort $s$ the set of index
  expressions $\{ e \sepbar \Delta \vdash e : s \}/\mathord\equiv$,
  quotiented by index expression equality. Index operations are
  interpreted by the corresponding syntactic operation on equivalence
  classes: $\semIndexExp{\texttt{\textup{f}}}([e_1],...,[e_n]) =
  [\texttt{\textup{f}}(e_1,...,e_n)]$.

  We take the relational interpretation of the primitive type
  $\tyPrimNm{X}$ as:
  \begin{displaymath}
    \semPrimType{\tyPrimNm{X}}(e) = \{ (x,x) \sepbar x \in \tySem{\tyPrimNm{X}} \land e \in \mathrm{Gen}(\{e_1,...,e_n\}) \}
  \end{displaymath}
  It is straightforward to check that for any index-erasure
  interpretation of the primitive operations $\eta_{\mathit{ops}} \in
  \Gamma_{\mathit{ops}}$, we have $(\eta_{\mathit{ops}},
  \eta_{\mathit{ops}}) \in \rsem{\Gamma_{\mathit{ops}}}*$ due to the
  fact that all members of $\Gamma_{\mathit{ops}}$ have well-generated
  types. Hence, by \thmref{thm:abstraction}, we know that for all
  terms (i.e. elements of the free model over $\Delta$)
  $e'_1\mathord:s_1,...,e'_m\mathord:s_m$:
  \begin{displaymath}%FIXME: check order of index environments
    \begin{array}{@{}l}
      \forall (x_1,x_1') \in \rsem{\tyPrim{X}{e_1}}(e'_1,...,e'_m), ..., \\
      \quad(x_n, x'_n) \in \rsem{\tyPrim{X}{e_n}}(e'_1,...,e'_m). \\
      \quad\quad (\tmSem{M} \eta_{\mathit{ops}}~x_1 ~...~x_n, \tmSem{M}\eta_{\mathit{ops}}~x'_1~...~x'_n) \in \rsem{\tyPrim{X}{e}}(e'_1,...,e'_m)
    \end{array}
  \end{displaymath}
  By setting $e'_j = i_j$, and using an arbitrary element $x \in
  \tySem{\tyPrimNm{X}}$ (which we have assumed non-empty), we have,
  for all $k$, $(x,x) \in \rsem{\tyPrim{X}{e_k}}(i_1,...,i_m)$ since
  $e_k$ is a member of the set of terms we are using to generate
  terms. Hence, $(\tmSem{M} \eta_{\mathit{ops}}~x...x,
  \tmSem{M}\eta_{\mathit{ops}}~x...x) \in \rsem{\tyPrimNm{X}}(e)$ and
  so $e \in \Gen(\{e_1,...,e_n\})$.
\end{proof}

% When applying \thmref{thm:general-nondef}, it is useful to observe
% that the free model constructed from the index variables $i_1,...,i_m$
% often has a well-known structure. Moreover, the set $\Gen(S)$ is
% always a sub-model of the free model. These observations can make it
% easy to apply the theorem in specific cases.

\paragraph{Application to \exref{ex:uninhabited-type}}
\thmref{thm:general-nondef} can be directly applied to show that the
type $\forall t\mathord:\SynTransl{2}.~\tyPrim{vec}{t + t} \to
\tyPrim{vec}{t}$ has no inhabitants. The free model over the single
index variable $t$ is (isomorphic to) the integers, and the sub-model
generated by the index expression $t + t$ corresponds to the even
integers. The result now follows simply because $1$ (i.e. the
interpreration of $t$) is not an even number.

\paragraph{Abelian Group Indexed Types} Kennedy
\cite{kennedy97relational} gave a general characterisation of
definability at first-order in the case of abelian group indexing in
terms of integer solutions to a set of linear equations. Specialising
\thmref{thm:general-nondef} to the case of abelian group indexing
yields Kennedy's characterisation.

\paragraph{Polymorphic Constants} \thmref{thm:general-nondef} does not
apply in the case when we have polymorphic constants. This is the case
with the polymorphic $0 : \forall
s\mathord:\SynGL{1}.~\tyPrim{real}{s}$ in
\autoref{fig:real-ops}. \thmref{thm:general-nondef} does not apply
because the index expression $s$ is not generated by the empty set:
$0$'s type is not well-generated. Nevertheless, it is easy to adapt
the proof of \thmref{thm:general-nondef} to handle a polymorphic
constant like $0$ by setting the relational interpretation of
$\tyPrimNm{X}$ to be:
\begin{displaymath}
  \rsem{X}(e) = \{ (x,x) \sepbar x \in \tySem{\tyPrimNm{X}} \land (x = 0 \lor e \in \Gen(\{e_1,...,e_n\})) \}
\end{displaymath}
The conclusion of the theorem now states that either we have $e \in
\Gen(\{e_1,...,e_n\})$ \emph{or} $\tmSem{M}$ is the constant $0$
function. This extended theorem can now be use to recreate the proof
given above in \exref{ex:sqrt-root-nondef}.

\paragraph{Adding Index Operations} \thmref{thm:general-nondef} also
does not directly apply in the case of \exref{ex:cube-root-nondef},
again because assumption that the types of the primitive operations
are all well-generated is not satisfied. In this case, the assumed
square root operation has type $\forall
s\mathord:\SynGL{1}.~\tyPrim{real}{s^2} \to \tyPrim{real}{s}$, and as
we observed at the start of this section, $s$ is not in the set
generated by $s^2$. However, to enable the application of the theorem,
we can assume an additional index operation $-^{1/2}$, acting
like square root at the index level. Now the free model over a single
generator is exactly the dyadic rationals, and the generated sub-model
consists of the dyadic rationals of the form $\frac{3k}{2^n}$.

%%% Local Variables:
%%% TeX-master: "paper"
%%% End: