### 2013-07-05 mechanics/paper.tex
Outline of Mechanics paper

\documentclass[preprint]{sigplanconf}

\newcommand{\sepbar}{\mathrel|}
\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{stmaryrd}

\title{From Parametricity to Conservation Laws, via Noether's Theorem}

\authorinfo{}{}{}
% \authorinfo{Robert Atkey}
%            {}
%            {bob.atkey@gmail.com}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{proposition}{Proposition}
\newcommand{\lemref}[1]{\hyperref[#1]{Lemma~\ref*{#1}}}
\newcommand{\thmref}[1]{\hyperref[#1]{Theorem~\ref*{#1}}}
\newcommand{\defref}[1]{\hyperref[#1]{Definition~\ref*{#1}}}
\newcommand{\propref}[1]{\hyperref[#1]{Proposition~\ref*{#1}}}
\newcommand{\corref}[1]{\hyperref[#1]{Corollary~\ref*{#1}}}
\newcommand{\conref}[1]{\hyperref[#1]{Conjecture~\ref*{#1}}}
\newcommand{\exref}[1]{\hyperref[#1]{Example~\ref*{#1}}}
\newcommand{\statementref}[1]{\hyperref[#1]{Statement~\ref*{#1}}}

\newcommand{\smoothrightarrow}{\stackrel{C^\infty}\longrightarrow}
\newcommand{\typeOfCartSp}[1]{\lbag #1 \rbag}
\newcommand{\elab}[1]{\lfloor #1 \rfloor}

\newtheoremstyle{examplestyle}
  {\topsep}  % space above
  {\topsep}  % space below
  {\normalfont}% name of font to use in the body of the theorem
  {0em}% measure of space to indent
  {\bfseries}% name of head font
  {.}% punctuation between head and body
  {5pt plus 1pt minus 1pt}% space after theorem head; " " = normal interword space
  {}% Manually specify head
\theoremstyle{examplestyle}
\newtheorem{example}{Example}
\newtheorem*{example*}{Example}

\newcommand{\fomega}{F$\omega$}
\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\Set}{\mathsf{Set}}
\newcommand{\Rel}{\mathsf{Rel}}
\newcommand{\Pow}{\mathcal{P}}
\newcommand{\frel}[1]{\langle #1 \rangle}
\newcommand{\coq}[1]{\textsf{#1}}
\newcommand{\CIC}{\textsc{CIC}}

\newcommand{\semKU}[1]{\llbracket #1 \rrbracket^O}
\newcommand{\semKR}[1]{\llbracket #1 \rrbracket^R}
\newcommand{\semKI}[1]{\llbracket #1 \rrbracket^{\mathrm{id}}}

\begin{document}

\maketitle

\begin{abstract}
  Invariance is of paramount importance in programming languages and
  in physics. In programming languages, John Reynolds' theory of
  relational parametricity demonstrates that parametric polymorphic
  programs are invariant under change of data representation, a
  property that yields ``free'' theorems about programs just from
  their types. In physics, Emmy Noether showed that if the action of a
  physical system is invariant under change of coordinates, then the
  physical system has a conserved quantity: a quantity that remains
  constant for all time. Knowledge of conserved quantities can reveal
  deep properties of physical systems. For example, the conservation
  of energy, which by Noether's theorem is a consequence of a system's
  invariance under time-shifting.

  In this paper, we link Reynolds' relational parametricity with
  Noether's theorem for deriving conserved quantities. We propose an
  extension of System F$\omega$ with new kinds, types and term
  constants for writing programs that describe classical mechanical
  systems in terms of their Lagrangians. We show, by constructing a
  relationally parametric model of our extension of F$\omega$, that
  relational parametricity is enough to satisfy the hypotheses of
  Noether's theorem, and so to derive conserved quantities for free,
  directly from the polymorphic types of Lagrangians expressed in our
  system.
\end{abstract}

\section{Introduction}

\begin{enumerate}
\item Symmetry in programming and physics
  \begin{enumerate}
  \item In programming: replacement of one data type by another
  \item In physics: invariance under change of unobservable
    properties. FIXME: check this.
  \item What unifies these, at a conceptual level? Inability to
    distinguish between situations, given the tools provided. In both
    cases, this leads to interesting properties of such systems. In
    programming, we get ``free theorems''. In Physics, via Noether's
    theorem, we get ``conservation laws''.
  \item In this paper, we show that we can define a programming
    language for classical mechanics, based on System F$\omega$, where
    the abstraction theorem directly fulfils the hypothesis of
    Noether's theorem, allowing us to derive conservation laws.
  \end{enumerate}
\item Parametricity: polymorphic programs have symmetry properties
  \begin{enumerate}
  \item Example of a free theorem
  \item Example of a geometric free theorem (the area of a triangle
    example)
  \end{enumerate}
\item Noether's theorem: symmetry properties of descriptions of
  systems lead to conservation laws
\item Let's put the two together
  \begin{enumerate}
  \item Need invariance properties of smooth functions
  \item So we extend System F$\omega$ with 
  \end{enumerate}
\item We present the classical mechanics and Noether's theorem first,
  since it is likely to be more unfamiliar to readers.
\end{enumerate}

We restate the results of Atkey, Johann and Kennedy in the more
general setting of System F$\omega$... which is good because it allows
for a deeper interpretation of what the unit kinds are. Also, we can
introduce useful stuff like nat-indexed types without too much fuss.

\paragraph{Summary of the Paper}

\begin{itemize}
\item FIXME: Recall the necessary background for Noether's theorem
  (Section~\ref{sec:conservation-laws-from-symmetry}).
\item FIXME: Show how to get symmetry properties from types
  (Section~\ref{sec:symmetry-from-types}).
\item FIXME: In Section~\ref{sec:types-for-classical-mech}, we
  construct a type system for classical mechanics,
\item FIXME: Section~\ref{sec:examples} we present examples of using
  our extension of System F$\omega$ for writing invariant Lagrangians
  that describe physical systems.
\item FIXME: field theory
\end{itemize}

\section{Conservation Laws from Symmetry}
\label{sec:conservation-laws-from-symmetry}

Make sure to state up front the necessary background required. To
understand: basic(ish) calculus. We will be using the Calculus of
Variations, but only by stating the main results. We go for standard
physics notation, rather than anything more complicated.

\subsection{The Principle of Stationary Action}

Introduce by stating Newton's second law, and saying that it can be
derived from the principle of stationary action.

Analytical mechanics, Lagrangian and Hamiltonian mechanics. Basically
different mathematical reformulations of Newtonian mechanics that
yield new insights into the underlying structure of the kinds of
physical systems studied in classical mechanics.

\paragraph{Lagrangian and Action}

Lagrangian mechanics reformulates Newtonian mechanics in terms of

Two points: the Langrangian and generalised coordinates. Do
generalised coordinates as a secondary thing.

\begin{displaymath}
  \mathcal{S}[q;a;b] = \int_a^b L(t,q(t),\dot{q}(t)) \mathit{dt}
\end{displaymath}

\begin{example}
  A single particle under a potential field. Set the potential field
  to be (downward) gravity.
\end{example}

\paragraph{The Principle of Stationary Action}

The principle of stationary action

\begin{displaymath}
  \frac{d}{dt}\frac{\partial L}{\partial \dot{q_i}} - \frac{\partial L}{\partial q_i} = 0
\end{displaymath}

Explain the different derivatives going on here

Example: Newton's second law...

\subsection{Noether's Theorem} Noether's theorem provides us with deep
insights into the properties of the solutions of the Euler-Lagrange
equations. FIXME: motivate invariance.

\paragraph{Invariance of the Action} Let
\begin{displaymath}
  \mathcal{S}[q;a;b] = \int_a^b L(t,q(t),\dot{q}(t)) \mathit{dt}
\end{displaymath}
be the action of some physical system described by the Lagrangian
$L$. Assume a differentiable invertible function $\tau : \mathbb{R}
\to \mathbb{R}$ that transforms time in some way and a function
$\sigma : \mathbb{R} \times \mathbb{R}^n \to \mathbb{R}^n$ that
transforms the vector of generalised coordinates $\vec{q}$ at time $t$
to another vector of generalised coordinates $\sigma(t,\vec{q})$.

We say that the action $\mathcal{S}$ is invariant under the
transformations $\tau$ and $\sigma$ if it is the case that for all
$a$, $b$ and $q$ that:
\begin{displaymath}
  \int_a^b L(t,q,\dot{q}) \mathit{dt} = \int_{\tau(a)}^{\tau(b)} L(t', \sigma(\tau^{-1}(t'), q(t')), \mathbf{FIXME}) \mathit{dt'}
\end{displaymath}
FIXME: Remark on the use of $t$ and $t'$ in the integrals. FIXME:
state that all we are doing is transforming the time and the path by
the transformations $\tau$ and $\sigma$ respectively. State why it
will be useful for the

Remark on how to give a simpler version of the invariance principle by
applying some simple transformations of integrals (see notes).

\begin{example}
  Continue a simple example from above to show that its action is
  invariant under (e.g.) time translation.
\end{example}

\paragraph{Noether's Theorem} Noether's theorem applies to actions
that are \emph{continuously} invariant. FIXME: not quite right: write
out the maths first.

(a) parameterise $\tau$ and $\sigma$ by a real value $\epsilon$ such
that we get a continuous symmetry of the action; (b) state Noether's
theorem as we will use it.

\begin{theorem}
  Noether's theorem
\end{theorem}

\begin{example}
  Continue the simple example from above to show that the invariance
  under time translation yields conservation of energy.
\end{example}

State that we will do more examples later on, after we discuss how to
derive symmetry from types.

\section{Symmetry from Types}

The first part here will be at least partly taken from Atkey
2012. Work out a way of emphasising that we do not need to do anything
to extend the basic underlying semantic structures involved: reflexive
graphs are enough. This presentation differs from Atkey 2012 in that
we emphasise the reflexive graph structure more, in order to be able
to better expose the connection with the groupoid kinds in the section
below.

\subsection{Relational Parametricity for System F$\omega$}
\label{sec:refl-graphs-for-fomega}

Reflexive graphs give us a way of capturing the higher-dimensional
structure of types in a

Relational parametricity (Reynolds, BFSS), reflexive graphs
(Robinson/Rosolini, Hasegawa, me).

Explain the semantic setting. Explain the ``property'' thing.

\paragraph{Syntax of System F$\omega$} Present the basic syntax of the
kinds, types and terms of (predicative) System F$\omega$.

Small kinds. We will extend the kinds, types and terms later to
encompass the necessary geometrical constructs for classical
mechanics.

\begin{displaymath}
  \kappa ::= * \sepbar \kappa_1 \to \kappa_2 \sepbar \kappa_1 \times \kappa_2 \sepbar \cdots
\end{displaymath}

\begin{mathpar}
  \inferrule*
  {\kappa_1~\mathrm{small} \\ \kappa_2~\mathrm{small}}
  {\kappa_1 \to \kappa_2~\mathrm{small}}

  \inferrule*
  {\kappa_1~\mathrm{small} \\ \kappa_2~\mathrm{small}}
  {\kappa_1 \times \kappa_2~\mathrm{small}}
\end{mathpar}

Type equality... $\Theta \vdash A \equiv B : \kappa$

\begin{figure}[t]
  \centering
  \begin{mathpar}
    \inferrule*
    {\alpha : \kappa \in \Theta}
    {\Theta \vdash \alpha : \kappa}

    \inferrule*
    {\Theta, \alpha : \kappa_1 \vdash A : \kappa_2}
    {\Theta \vdash \lambda \alpha : \kappa_1.\ A : \kappa_1 \to \kappa_2}

    \inferrule*
    {\Theta \vdash F : \kappa_1 \to \kappa_2 \\ \Theta \vdash A : \kappa_1}
    {\Theta \vdash F A : \kappa_2}

    \inferrule*
    {\Theta \vdash A : \kappa_1 \\ \Theta \vdash B : \kappa_2}
    {\Theta \vdash \langle A, B \rangle : \kappa_1 \times \kappa_2}

    \inferrule* [right=${i \in \{1,2\}}$]
    {\Theta \vdash A : \kappa_1 \times \kappa_2}
    {\Theta \vdash \pi_i~A : \kappa_i}

    \inferrule*
    {\Theta \vdash A : * \\ \Theta \vdash B : *}
    {\Theta \vdash A \times B : *}

    \inferrule*
    {\Theta \vdash A : * \\ \Theta \vdash B : *}
    {\Theta \vdash A \to B : *}

    \inferrule*
    {\Theta, \alpha : \kappa \vdash A : * \\ \kappa\textrm{ small}}
    {\Theta \vdash \forall\alpha\mathord:\kappa.\ A : *}
  \end{mathpar}
  \caption{Types and their Kinds}
  \label{fig:types}
\end{figure}

\begin{figure}[t]
  \centering
  \begin{mathpar}
    \inferrule*
    {x : A \in \Gamma}
    {\Theta \sepbar \Gamma \vdash x : A}

    \inferrule*
    {\Theta \sepbar \Gamma \vdash e : A \\ \Theta \vdash A \equiv B : *}
    {\Theta \sepbar \Gamma \vdash e : B}
  \end{mathpar}
  \begin{mathpar}
    \inferrule*
    {\Theta \sepbar \Gamma \vdash e_1 : A_1 \\\\
      \Theta \sepbar \Gamma \vdash e_2 : A_2}
    {\Theta \sepbar \Gamma \vdash (e_1, e_2) : A_1 \times A_2}

    \inferrule* [right=${i \in \{1,2\}}$]
    {\Theta \sepbar \Gamma \vdash e : A_1 \times A_2}
    {\Theta \sepbar \Gamma \vdash \pi_i e : A_i}
  \end{mathpar}
  \begin{mathpar}
    \inferrule*
    {\Theta \sepbar \Gamma, x : A \vdash e : B}
    {\Theta \sepbar \Gamma \vdash \lambda x : A.\ e : A \to B}

    \inferrule*
    {\Theta \sepbar \Gamma \vdash e_1 : A \to B \\
      \Theta \sepbar \Gamma \vdash e_2 : A}
    {\Theta \sepbar \Gamma \vdash e_1 e_2 : B}

    \inferrule*
    {\Theta, \alpha : \kappa \sepbar \Gamma \vdash e : A \\ \alpha \not\in \mathit{fv}(\Gamma)}
    {\Theta \sepbar \Gamma \vdash \Lambda \alpha \mathord: \kappa.\ e : \forall \alpha \mathord:\kappa. A}

    \inferrule*
    {\Theta \sepbar \Gamma \vdash e : \forall \alpha\mathord:\kappa. A \\ \Theta \vdash B : \kappa}
    {\Theta \sepbar \Gamma \vdash e\ [B] : A\{B/\alpha\}}
  \end{mathpar}
  \caption{Terms and their Types}
  \label{fig:terms}
\end{figure}

\paragraph{Reflexive Graphs and the Interpretation of Kinds} We will
interpret every kind $\kappa$ is as a reflexive graph, which we now
define. A \emph{reflexive graph} is a triple $(O, R, \mathrm{id})$,
where $O$ is a large set of objects, $R : O \times O \to \Set$ assigns
a small set of directed `edges' to each pair of objects, and
$\mathrm{id} : \forall o \in O.~R(o,o)$ assigns a distinguished
`identity' edge from every object to itself. We think of the edges of
a reflexive graph as abstract ``relations'' between the
objects. Indeed, in the interpretation of the kind of types, $*$,
below, the edges will be exactly relations.

A \emph{small} reflexive graph is a reflexive graph $(O, R,
\mathrm{id})$ where $O$ is a small set of objects. We use small
reflexive graphs as the semantic interpretation of small kinds.

The interpretation of kinds as reflexive graphs, and small kinds as
small reflexive graphs is a key property of our semantics that we will
maintain as we add additional kinds in Sections XXX below.

\begin{property}\label{property:semantic-kinds}
  Each kind $\kappa$ is interpreted as a reflexive graph
  $\sem{\kappa}$. If $\kappa~\mathrm{small}$, then $\sem{\kappa}$ is a
  small reflexive graph.
\end{property}

An appealing interpretation of reflexive graphs is as ``categories
without composition''. Following this intuition, we define morphisms
of reflexive graphs as ``functors'', without the preservation of
composition condition. A \emph{morphism of reflexive graphs}
$(O_1,R_1,\mathrm{id}_1)$ and $(O_2, R_2, \mathrm{id}_2)$ is a pair of
mappings $f : O_1 \to O_2$ and $r : \forall o, o' \in
O_1.~R_1(o,o') \to R_2(f~o, f~o')$ such that identities are
preserved: $r~o~o~(\mathrm{id}_1~o) = \mathrm{id}_2~(f~o)$. We use
morphisms of reflexive graphs below to interpret well-kinded types.

We will use the notation $-^O$, $-^R$ and $-^{\mathrm{id}}$ for the
first, second and third projections out of tuples representing
reflexive graphs. Similarly, we use $-^f$ and $-^r$ for the first and
second projections out of tuples representing reflexive graph
morphisms.

We now define the interpretations of the basic kinds of System
F$\omega$ we defined above, making sure that we maintain
Property~\ref{property:semantic-kinds}. At base kind, the collection
of objects is simply the type of all small sets; edges between $A$ and
$B$ are binary relations on $A$ and $B$ (i.e., subsets of $A \times
B$); and the distinguished identity edge is exactly the equality
relation:
\begin{displaymath}
  \sem{*} = (\Set, \Rel, \equiv)
\end{displaymath}
The reflexive graph $\sem{*}$ is not small, due to the collection of
all small sets $\Set$ not forming a small set.

For higher kinds $\kappa_1 \to \kappa_2$, the collection of objects
consists of reflexive graph morphisms from the interpretation of
$\kappa_1$ to the interpretation of $\kappa_2$; the edges between
morphisms $(f, r)$ and $(f', r')$ are edge transformers; and the
distinguished identity relation for $(f,r)$ is just $r$:
\begin{displaymath}
  \begin{array}{@{}l}
    \sem{\kappa_1 \to \kappa_2} = \\
    \begin{array}[t]{l@{}l@{}}
      (& \{ (f,r) \sepbar (f, r) : \sem{\kappa_1} \to \sem{\kappa_2} \},\\
      &((f,r),(f',r')) \mapsto \forall o,o'. \semKR{\kappa_1}(o,o') \to \semKR{\kappa_2}(f~o, f'~o'), \\
      &(f,r) \mapsto r\ )
    \end{array}
  \end{array}
\end{displaymath}
By the assumption that our collection of small sets is closed under
the formation of function spaces and set comprehension, if
$\sem{\kappa_1}$ and $\sem{\kappa_2}$ are small reflexive graphs, then
so is $\sem{\kappa_1 \to \kappa_2}$.

Product kinds are interpreted by taking the product of their
interpretations as reflexive graphs:
\begin{displaymath}
  \begin{array}{@{}l}
    \sem{\kappa_1 \times \kappa_2} = \\
    \begin{array}[t]{l@{}l}
      (& \semKU{\kappa_1} \times \semKU{\kappa_2}, \\
      & ((o_1,o_2),(o_1',o_2')) \mapsto \semKR{\kappa_1}(o_1,o_1') \times \semKR{\kappa_2}(o_2,o_2'), \\
      & (o_1,o_2) \mapsto (\semKI{\kappa_1}(o_1), \semKI{\kappa_2}(o_2)))
    \end{array}
  \end{array}
\end{displaymath}
The objects of the interpretation of a product kind is the product of
the underlying collections of objects of the two parts, and the
relational component is simply the product of the relational
components. This naturally leads to the identity component being
defined as the tuple of the identity components of the two
parts. Again, by assumption that our collection of small sets is
closed under products, if $\sem{\kappa_1}$ and $\sem{\kappa_2}$ are
small reflexive graphs, then so is $\sem{\kappa_1 \times
  \kappa_2}$. Product kinds generalise to the interpretation of
kinding contexts $\Theta = \alpha_1 : \kappa_1, ..., \alpha_n :
\kappa_n$, which are interpreted as the product of the reflexive graph
interpretations of $\kappa_1, ..., \kappa_n$:
\begin{displaymath}
  \begin{array}{l}
    \sem{\alpha_1 : \kappa_1, ..., \alpha_n : \kappa_n} = \\
    \begin{array}[t]{l@{}l}
      (&\semKU{\kappa_1} \times ... \times \semKU{\kappa_n}, \\
      &(\theta,\theta') \mapsto \semKR{\kappa_1}(\pi_1\theta, \pi_1\theta') \times ... \times \semKR{\kappa_n}(\pi_n\theta, \pi_n\theta'), \\
      &\theta \mapsto (\semKI{\kappa_1}(\pi_1 \theta), ..., \semKI{\kappa_n}(\pi_n \theta)))
    \end{array}
  \end{array}
\end{displaymath}

In Sections~\ref{sec:discrete-kinds} and~\ref{sec:groupoid-kinds}
below, we will extend System F$\omega$ with additional kinds, and
assign them reflexive graph interpretations, making sure that we
maintain Property~\ref{property:semantic-kinds}.

\begin{figure*}[t]
  \begin{displaymath}
    \begin{array}{l@{\hspace{0.3em}}c@{\hspace{0.3em}}l}
      \sem{\Theta \vdash \alpha_i : \kappa_i}^f \theta & = & \pi_i \theta \\
      \sem{\Theta \vdash \alpha_i : \kappa_i}^r\theta\theta' \rho & = & \pi_i \rho \\
      \\
      \sem{\Theta \vdash \lambda \alpha \mathord: \kappa_1.~A : \kappa_1 \to \kappa_2}^f \theta & = & (
      \begin{array}[t]{@{}l}
        \lambda o \in \semKU{\kappa_1}.\ \sem{A}^f(\theta, o), 
        \lambda o, o' \in \semKU{\kappa_1}, r \in \semKR{\kappa_1}(o,o').\ \sem{A}^r(\theta,o)(\theta,o')(\semKI{\Theta}\theta, r))
      \end{array}\\
      \sem{\Theta \vdash \lambda \alpha \mathord: \kappa_1.~A : \kappa_1 \to \kappa_2}^r \theta\theta' \rho & = &
      \lambda o,o' \in \semKU{\kappa_1}, r \in \semKR{\kappa_1}(o,o').\ \sem{A}^r(\theta,o)(\theta',o')(\rho, r) \\
      \\
      \sem{\Theta \vdash F A : \kappa_2}^f \theta & = & \pi_1 (\sem{F}^f\ \theta)\ (\sem{A}^f\ \theta)\\
      \sem{\Theta \vdash F A : \kappa_2}^r\theta\theta'\rho & = & \sem{F}^r\theta\theta'\rho\ (\sem{A}^f \theta)(\sem{A}^f \theta')(\sem{A}^r\theta\theta'\rho) \\
      \\
      \sem{\Theta \vdash \langle A, B \rangle : \kappa_1 \times \kappa_2}^f \theta & = & (\sem{A}^f \theta, \sem{B}^f \theta) \\
      \sem{\Theta \vdash \langle A, B \rangle : \kappa_1 \times \kappa_2}^r \theta \theta' \rho & = & (\sem{A}^r \theta \theta' \rho, \sem{B}^r \theta \theta' \rho) \\
      \\
      \sem{\Theta \vdash \pi_i A : \kappa_i}^f \theta & = & \pi_i (\sem{A}^f \theta) \\
      \sem{\Theta \vdash \pi_i A : \kappa_i}^r \theta \theta' \rho & = & \pi_i (\sem{A}^r \theta \theta' \rho)
    \end{array}
  \end{displaymath}
  \caption{Interpretation of type-level $\lambda$-calculus as reflexive graph morphisms}
  \label{fig:type-level-lambda-calc}
\end{figure*}

\begin{figure*}[t]
  \begin{displaymath}
    \begin{array}{l@{\hspace{0.3em}}c@{\hspace{0.3em}}l}
      \sem{\Theta \vdash A \to B : *}^f \theta & = & \sem{A}^f\theta \to \sem{B}^f\theta \\
      \sem{\Theta \vdash A \to B : *}^r\theta\theta'\rho & = &
      \{ (f,f') \sepbar \forall (a,a') \in \sem{A}^r\theta\theta'\rho.\ (f~a, f'~a') \in \sem{B}^r\theta\theta'\rho \} \\
      \\
      \sem{\Theta \vdash A \times B : *}^f \theta & = & \sem{A}^f\theta \times \sem{B}^f\theta \\
      \sem{\Theta \vdash A \times B : *}^r\theta\theta'\rho & = &
      \{ ((a,b),(a',b')) \sepbar (a,a') \in \sem{A}^r\theta\theta'\rho, (b,b') \in \sem{B}^r\theta\theta'\rho \} \\
      \\
      \sem{\Theta \vdash \forall \alpha\mathord:\kappa.~A : *}^f \theta &=& \{
      \begin{array}[t]{@{}l}
        x \in (\forall o \in \semKU{\kappa}.\ \sem{A}^f(\theta, o)) \sepbar
        \forall o,o',r \in \semKR{\kappa}(o,o').\ (x~o, x~o') \in \sem{A}^r(\theta,o)(\theta,o')(\semKI{\Theta} \theta, r) \}
      \end{array} \\
      \sem{\Theta \vdash \forall \alpha\mathord:\kappa.~A : *}^r\theta\theta'\rho &=& \{ (x,x') \sepbar \forall o, o', r \in \semKR{\kappa}(o,o').\ (x~o, x'~o') \in \sem{A}^r(\theta,o)(\theta',o')(\rho, r) \}
    \end{array}
  \end{displaymath}
  \caption{Interpretation of basic types as reflexive graph morphisms}
  \label{fig:basic-type-interpretation}
\end{figure*}

\paragraph{Interpretation of Types} Well-kinded types $\Theta \vdash A
: \kappa$ from Figure~\ref{fig:types} are interpreted as reflexive
graph morphisms $\sem{A} : \sem{\Theta} \to \sem{\kappa}$. We sum this
up as a property of our semantics:
\begin{property}\label{property:semantic-types}
  Each well-kinded type $\Theta \vdash A : \kappa$ is interpreted as a
  reflexive graph morphism $\sem{A} : \sem{\Theta} \to \sem{\kappa}$,
  such that if $\Theta \vdash A \equiv B : \kappa$, then $\sem{A} =
  \sem{B}$.
\end{property}
%
% That is, every well-kinded type is interpreted as a pair $(\sem{A}^f,
% \sem{A}^r)$ of functions $\sem{A}^f : \sem{\Theta}^O \to
% \sem{\kappa}^O$ and $\sem{A}^r : \forall o, o'.~\sem{\Theta}^R(o,o')
% \to \sem{\kappa}^R(\sem{A}^f~o, \sem{A}^f~o')$, such that identity
% edges are preserved.
%
The interpretation of the $\lambda$-calculus fragment (i.e.,
variables, $\lambda$-abstraction and application, and products) of the
language of well-kinded types is displayed in
Figure~\ref{fig:type-level-lambda-calc}. The interpretations are
unsurprising given the reflexive graph interpretation of the kinds
$\kappa_1 \to \kappa_2$ and $\kappa_1 \times \kappa_2$ we gave above.

Figure~\ref{fig:basic-type-interpretation} shows the interpretations
of the basic type constructors for function and product types, and
universal quantification. Each of these constructs builds an object of
kind $*$, so the object-level interpretation is a small set in $\Set$,
and the relation-level interpretation is an actual relation. In the
cases of the function and product types, the object-level
interpretation is just as set-theoretic function and product
respectively, and the relation-level interpretation uses the standard
logical relations interpretations of these type
constructors. Universal quantification, $\forall
\alpha\mathord:\kappa.~A$, is interpreted at the object level by
taking the dependent product over the objects of the interpretation of
$\kappa$ (this product exists because we have stipulated that $\kappa$
must be small), and then restricting to those elements of the
dependent product that preserve relations. This restriction is
required for this interpretation to preserve identity edges, and so be
a reflexive graph morphism. The relation-level interpretation of
universal quantification is the standard relational interpretation of
such types, albeit here generalised to kinds interpreted as arbitrary
reflexive graphs.

Well-kinded typing contexts $\Theta \vdash \Gamma$ are interpreted as
reflexive graph morphisms $\sem{\Gamma} : \sem{\Theta} \to \sem{*}$ by
taking the product of the interpretations of their constituent types,
similar to the interpretation of the product types $A \times B$.

We will extend basic System F$\omega$ with additional types and type
equalities in Sections~\ref{sec:discrete-kinds},
\ref{sec:groupoid-kinds}, and~\ref{sec:types-for-classical-mech},
below. These new types will also be assigned interpretations as
reflexive graph morphisms, and we will ensure that
Property~\ref{property:semantic-types} is maintained.

\paragraph{Interpretation of Terms} We omit the straightforward and
relatively uninteresting interpretation of well-typed terms $\Theta
\sepbar \Gamma \vdash e : A$, and just state that there is a
well-defined function interpreting each well-typed term, with the
property that it takes related environments to related results:

\begin{property}
  For all well-typed terms $\Theta \sepbar \Gamma \vdash e : A$ there
  is a function $\sem{e} \in (\forall \theta \in \semKU{\Theta}.\
  \sem{\Gamma}^f\theta \to \sem{A}^f\theta)$, such that, for all
  $\theta, \theta' \in \semKU{\Theta}$, $\rho \in
  \semKR{\Theta}(\theta,\theta')$, $\gamma \in \sem{\Gamma}^f\theta$
  and $\gamma' \in \sem{\Gamma}^f\theta'$, if $(\gamma, \gamma') \in
  \sem{\Gamma}^r\theta\theta'\rho$ then $(\sem{e}\theta\gamma,
  \sem{e}\theta'\gamma') \in \sem{A}^r\theta\theta'\rho$.  Moreover,
  this interpretation is sound for the $\beta\eta$ equational theory
  of terms.
\end{property}

\subsection{Discrete Kinds}
\label{sec:discrete-kinds}

In Section~\ref{sec:refl-graphs-for-fomega}, we only had a single base
kind: the kind $*$ of proper types.

Natural numbers as a discrete kind. The type of length-indexed
vectors.

\begin{displaymath}
  \kappa ::= \cdots \sepbar \mathsf{nat}
\end{displaymath}

\begin{displaymath}
  \inferrule*
  { }
  {\mathsf{nat}~\mathrm{small}}
\end{displaymath}

\begin{displaymath}
  \sem{\mathsf{nat}} = (\mathbb{N}, (n_1,n_2) \mapsto \{ * \sepbar n_1 = n_2 \}, n \mapsto *)
\end{displaymath}

Make the point that any set can be used as the interpretation of a
kind. Also, the utility of singleton types.

\subsection{Group(oid) Kinds}
\label{sec:groupoid-kinds}

There is a full and faithful functor from the category of groupoids to
the category of reflexive graphs, that preserves smallness.

\begin{enumerate}
\item Each (matrix) group gives a kind, which we can use as parameter
  to a type.
\item The collection of cartesian spaces, with diffeomorphisms as the
  edges.
\item Do units of measure as an example (include one of the results
  from that paper, e.g., type iso)
\item Mention groupoid kinds, and link to Hofmann and Streicher. Do
  finite sets and bijections as an example.
\end{enumerate}

\paragraph{Why not Groupoids Everywhere?} Explain why we aren't using
groupoids as kinds: don't need to, doesn't directly generalise
relational parametricity, see Robinson's paper. Distance-indexed
types.

\section{A Type System for Classical Mechanics}
\label{sec:types-for-classical-mech}

FIXME: rewrite this para to refer back concretely

We have a way of getting symmetry properties from types, and a way of
using symmetry properties to derive conservation laws for physical
systems. We now put the two together by building an applied version
System F$\omega$ with new kind-, type- and term-level constants that
will deliver us the exact symmetry properties we require to be able to
apply Noether's theorem.

\subsection{Extending System F$\omega$ for Classical Mechanics}

Let us enumerate our requirements for an extension of System F$\omega$
for writing invariant Lagrangians\footnote{Make sure that I say this
  is the plan back in Sec 2}.

\begin{enumerate}
\item Kind of cartesian spaces, parameterised by a matrix group and a
  translation group
\item Smooth functions between spaces: motivate this by looking at the
  requirements of the Euler-Lagrange equations.
\item Primitives for dealing with the geometry of the spaces (inner
  product, vector space operations, cross product(?), trigonometric
  functions (at least sin and cos)).\footnote{FIXME: fix the notation of $\forall$}
  \begin{displaymath}
    \begin{array}{l@{\hspace{0.3em}}c@{\hspace{0.3em}}l}
      \sin &:& \forall z \mathord: \mathsf{Z}.~\mathbb{R}\langle 2\pi \cdot z, 1 \rangle \smoothrightarrow \mathbb{R} \langle 0, 1 \rangle \\
      \cos &:& \forall z \mathord: \mathsf{Z}.~\mathbb{R}\langle 2\pi \cdot z, 1 \rangle \smoothrightarrow \mathbb{R} \langle 0, 1 \rangle \\
    \end{array}
  \end{displaymath}
\end{enumerate}

\begin{displaymath}
  \begin{array}{l@{\hspace{0.3em}}c@{\hspace{0.3em}}l}
    \mathrm{exp} &:& \forall t \mathord: \mathsf{T}(1).~\mathbb{R}\langle t, 1 \rangle \smoothrightarrow \mathbb{R}\langle 0, \mathit{exp}~t \rangle
  \end{array}
\end{displaymath}

Group homomorphisms:
\begin{itemize}
\item $\mathit{exp} : \mathsf{T}(1) \to \mathsf{GL}(1)$
\item $\mathit{scale}_n : \mathsf{GL}(1) \to \mathsf{GL}(n)$
\item $\mathit{ortho}_n : \mathsf{O}(n) \to \mathsf{GL}(n)$
\item $\underline{c}\cdot- : \mathsf{Z} \to \mathsf{T}(1)$
\end{itemize}
Group operations on translations (for all $n$), and on
$\mathsf{GL}(1)$ (no abs, because it is not smooth).

\begin{displaymath}
  \kappa ::= \cdots \sepbar \mathsf{CartSp} \sepbar \mathsf{T}(n) \sepbar \mathsf{GL}(n) \sepbar \mathsf{O}(n) \sepbar \mathsf{Z} \sepbar XXX
\end{displaymath}

\begin{displaymath}
  \begin{array}{l@{\hspace{0.3em}}c@{\hspace{0.3em}}l}
    (\smoothrightarrow) &:& \mathsf{CartSp} \to \mathsf{CartSp} \to * \\
    \mathbb{R}^n &:& \mathsf{T}(n) \to \mathsf{GL}(n) \to \mathsf{CartSp} \\
    (\times) &:& \mathsf{CartSp} \to \mathsf{CartSp} \to \mathsf{CartSp}
  \end{array}
\end{displaymath}

\begin{itemize}
\item $\mathsf{CartSp}$ is the kind of cartesian spaces
  ($\mathbb{R}^n$, for some $n$) and diffeomorphisms between them,
  with the standard definition of diffeomorphism on these spaces. The
  identity edge is just the identity function.
\item $\mathbb{R}^n~t~g$ is the cartesian space with the
  diffeomorphism defined by the action of the translation and matrix
  group.
\item $X \times Y$ takes the product of two cartesian spaces, and the
  corresponding product of the diffeomorphisms.
\item $X \smoothrightarrow Y$ consists of the infinitely
  differentiable functions from $X$ to $Y$ that preserve the
  diffeomorphisms.
\end{itemize}

\begin{enumerate}
\item Need coercions between the different groups, as required by the
  examples.
\item Need operations on vectors, e.g., the usual linear space ones,
  and inner product. Perhaps also the cross product one.
\item What operations on the smooth function space do we need?
  composition, products, identities.
\end{enumerate}

\begin{figure*}[t]
  \centering
  \textbf{Administrative rules}
  \begin{mathpar}
    \inferrule*
    {x : \typeOfCartSp{X} \in \Gamma}
    {\Theta \sepbar \Gamma; \Delta \vdash x : X}

    \inferrule*
    {x : X \in \Delta}
    {\Theta \sepbar \Gamma; \Delta \vdash x : X}

    \inferrule*
    {\Theta \sepbar \Gamma; \Delta \vdash e_1 : X \\
      \Theta \sepbar \Gamma; \Delta, x : X \vdash e_2 : Y}
    {\Theta \sepbar \Gamma; \Delta \vdash \mathrm{let}~x = e_1~\mathrm{in}~e_2 : Y}

    \inferrule*
    {\Theta \sepbar \Gamma \vdash e_1 : C^\infty(X, Y) \\\\
      \Theta \sepbar \Gamma; \Delta \vdash e_2 : X}
    {\Theta \sepbar \Gamma; \Delta \vdash e_1 (e_2) : Y}
  \end{mathpar}
  \textbf{Vector space operations, and dot product}
  \begin{mathpar}
    \inferrule*
    {\vec{x} \in \mathbb{R}^n}
    {\Theta \sepbar \Gamma; \Delta \vdash \vec{x} : \mathbb{R}^n\langle 1, 0 \rangle}

    \inferrule*
    {\Theta \vdash g : \mathsf{GL}(n)}
    {\Theta \sepbar \Gamma; \Delta \vdash 0 : \mathbb{R}^n\langle g, 0 \rangle}

    \inferrule*
    {\Theta \sepbar \Gamma; \Delta \vdash e_1 : \mathbb{R}^n\langle g, t_1 \rangle \\
      \Theta \sepbar \Gamma; \Delta \vdash e_2 : \mathbb{R}^n\langle g, t_2 \rangle}
    {\Theta \sepbar \Gamma; \Delta \vdash e_1 + e_2 : \mathbb{R}^n\langle g, t_1 + t_2 \rangle}

    \inferrule*
    {\Theta \sepbar \Gamma; \Delta \vdash e_1 : \mathbb{R}^n\langle g, t_1 \rangle \\
      \Theta \sepbar \Gamma; \Delta \vdash e_2 : \mathbb{R}^n\langle g, t_2 \rangle}
    {\Theta \sepbar \Gamma; \Delta \vdash e_1 - e_2 : \mathbb{R}^n\langle g, t_1 - t_2 \rangle}

    \inferrule*
    {\Theta \sepbar \Gamma; \Delta \vdash e_1 : \mathbb{R}\langle g_1, 0 \rangle \\
      \Theta \sepbar \Gamma; \Delta \vdash e_2 : \mathbb{R}^n\langle g_2, 0 \rangle}
    {\Theta \sepbar \Gamma; \Delta \vdash e_1 * e_2 : \mathbb{R}^n\langle \mathit{scale}_n(g_1)g_2, 0 \rangle}

    \inferrule*
    {\Theta \sepbar \Gamma; \Delta \vdash e_1 : \mathbb{R}^n\langle \mathit{scale}_n(s)\mathit{ortho}_n(o), 0 \rangle \\
      \Theta \sepbar \Gamma; \Delta \vdash e_2 : \mathbb{R}^n\langle \mathit{scale}_n(s)\mathit{ortho}_n(o), 0 \rangle}
    {\Theta \sepbar \Gamma; \Delta \vdash e_1 \cdot e_2 : \mathbb{R}\langle\mathit{scale}_n(s)^2, 0 \rangle}
  \end{mathpar}
  \textbf{Transcendental functions}
  \begin{mathpar}
    \inferrule*
    {\Theta \sepbar \Gamma; \Delta \vdash e : \mathbb{R}\langle 1, t \rangle}
    {\Theta \sepbar \Gamma; \Delta \vdash \mathrm{exp}~e : \mathbb{R}\langle \mathit{exp}~t, 0 \rangle}

    \inferrule*
    {\Theta \sepbar \Gamma; \Delta \vdash e : \mathbb{R}\langle 2\pi z, 0 \rangle}
    {\Theta \sepbar \Gamma; \Delta \vdash \mathrm{sin}~e : \mathbb{R}\langle 1, 0 \rangle}

    \inferrule*
    {\Theta \sepbar \Gamma; \Delta \vdash e : \mathbb{R}\langle 2\pi z, 0 \rangle}
    {\Theta \sepbar \Gamma; \Delta \vdash \mathrm{cos}~e : \mathbb{R}\langle 1, 0 \rangle}
  \end{mathpar}
  FIXME: products, division, n-ary products (sum, zip, reduce), constants
  \caption{Type system for smooth functions}
  \label{fig:smooth-terms}
\end{figure*}

\begin{theorem}
  If $\Theta \sepbar \Gamma; \Delta \vdash e : X$ then $\Theta \sepbar
  \Gamma \vdash \elab{e} : C^\infty(\elab{\Delta}, X)$.
\end{theorem}

\section{Examples}

$- \sepbar V : \forall ... . C^\infty(\mathbb{R}^3\langle o, t \rangle, \mathbb{R}\langle 1, 0 \rangle); t : \mathbb{R}\langle 1, t_t \rangle, q : \mathbb{R}^3\langle o, t_q \rangle, \dot{q} : \mathbb{R}^3\langle o, 0 \rangle \vdash \frac{1}{2} * m * (\dot{q} \cdot \dot{q}) - V(q) : \mathbb{R}\langle 1, 0 \rangle$

\begin{enumerate}
\item free particle
\item $2$-body
\item $n$-body (use length-indexed vectors)
\item pendulums
\item oscillators
\end{enumerate}

\section{Basic Field Theory}

\subsection{Principle of Stationary Action for Fields}

\subsection{Noether's Theorem for Fields}

\subsection{The Klein-Gordon Equation}

\section{Conclusions}

\begin{enumerate}
\item Gauge theory
\item Quantum field theory
\item Dependent types
\item Homotopy type theory
\item Distance indexed types and differential privacy.
\item Languages with function spaces with more structure (Alex's
  Enriched Effect Calculus) (Idea for differential privacy)
\end{enumerate}

\newpage

Overall plan:
\begin{enumerate}
\item A type system for writing smooth functions between smooth
  manifolds, subject to invariance with respect to certain Lie groups
\item Use this to write Lagrangians that describe classical mechanics
  situations, in such a way that Noether's theorem can be applied.
\item Need to explain:
  \begin{enumerate}
  \item Lagrangians
  \item Euler-Lagrange equations for $L(t,q,q')$
    \begin{displaymath}
      \frac{\mathrm{d}}{\mathrm{d}t} \frac{\partial L}{\partial q'} - \frac{\partial L}{\partial q} = 0
    \end{displaymath}
  \item Lie groups (manifolds!?), and their infinitesimal generators
    (see the exposition in Olver's book).
  \item Noether's theorem, with (lots of) examples. Explain what was
    in Noether's original paper (first theorem vs. second theorem).
  \end{enumerate}
  Interleave the presentation of the maths/physics and the type
  system?
\item Stick to being relatively uncomplicated with respect to the
  manifold stuff? Or use dependent types to handle the tangent spaces
  ``properly''?
\item Operations we want (in order to do the examples): distance
  between points (arclength/geodesic length), trigonometric functions
  (what continuous invariance properties do these have?),
\item Possibly use a variant of System F$\omega$ as the
  superstructure: System F$\omega$ plus a sub-language for dealing
  with smooth functions (which, in principle, could be handled just by
  using combinators to build something in a arrow-style). System
  F$\omega$ gives us a lot of power, and means that we don't need to
  go anywhere near dependent types...
  \begin{enumerate}
  \item So manifolds are a kind
  \item And so are Lie groups (?), not really, they have a different
    status (due to the identity extension requirement)
  \item Flat spaces are type operators from groups that act on them to
    manifolds
  \item Have a type constructor that builds the space of infinitely
    differentiable functions between two spaces
  \item Elaboration process that desugars a nicer syntax for smooth
    functions...
  \item Possibly a bit of natural numbers and vectors to allow for
    $n$-body problems...
  \item Need to restrict to predicativity in order to get classical
    mathematics? (probably should just forget about type
    quantification, and restrict the set of things we can quantify
    over to certain ``small'' kinds; just the groups?). Mention that
    going for SDG as a basis might offer a way out of this.
  \item Sub-kinding?
  \end{enumerate}
\item Examples:
  \begin{enumerate}
  \item $n$-body problem in $\mathbb{R}^3$, with all the consequent
    conservation laws.
  \item Central force problem
  \item Pendulums
  \item Rigid body problem? Use quotients in the definitions of the
    trigonometric functions? $\sin : \forall n :
    \mathbb{Z}.~\mathbb{R}\langle2\pi n\rangle \to \mathbb{R}\langle 1
    \rangle$. This gives us a way of doing circles in a neat way.
  \end{enumerate}
\end{enumerate}

\begin{enumerate}
\item Be clear that we don't really expect these programs to
  necessarily be (directly) executable.
\item Manifolds?
  \begin{enumerate}
  \item Spheres
  \item Tangent spaces
  \item 'Flat' spaces
  \item (Quotient spaces?) (have a look at Olver's book)
  \end{enumerate}
\end{enumerate}

Groups and their operations:
\begin{enumerate}
\item General linear group
\item Orthogonal group
\item Rotations around certain axes (in 
\end{enumerate}

Types (maybe include dependent types, for $n$-ary sized problems?)
\begin{displaymath}
  \begin{array}{rcl}
    M,M' & ::= & \mathbb{R}^n\langle e \rangle \sepbar \Sigma x : M.~M' \sepbar T_xM \\
    A, B & ::= & S \sepbar C^\infty(S,T) \sepbar A \times B \sepbar \forall G.~A \sepbar A \to B \sepbar A + B
  \end{array}
\end{displaymath}

\begin{mathpar}
  \inferrule*
  {\Gamma \sepbar x : S \vdash M :^\infty T}
  {\Gamma \vdash \lambda^\infty x : S.~M : C^\infty(S,T)}
\end{mathpar}

So a Lagrangian is something of type:
\begin{displaymath}
  \forall G:\mathsf{E}(3).~C^\infty(\mathbb{R} \times \mathbb{R}^3\langle G \rangle \times \mathbb{R}^3\langle G \rangle, \mathbb{R})
\end{displaymath}

\section{Introduction}

\section{Lagrangian Mechanics}

\section{Symmetry and Conservation Laws}

\section{A Type System for Lagrangians}

Syntax and Semantics, and the abstraction theorem...

\section{Examples}

\section{Field Theory}

\end{document}